### 2009-12-09 paper/paper.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@364 3fc257bb-8550-0410-81b3-a5c103da83f7

\documentclass[11pt]{article}
\usepackage[pdftex]{graphicx}

\setlength{\textwidth}{6.5in}
\setlength{\textheight}{9.25in}

\setlength{\voffset}{-0.0in}
\hoffset=-50pt
\topmargin=0pt
\headheight=0pt
\headsep=0pt
\evensidemargin=72pt % 1in+7pt

\setlength{\parskip}{0.5ex}
\setlength{\parindent}{1em}
\setlength{\floatsep}{1pt}
\setlength{\textfloatsep}{2pt}
\setlength{\intextsep}{2pt}

\begin{document}

\title{Map{R}educe in {MPI} for Large-scale Graph Algorithms}

\author{
Steven J.~Plimpton, Karen D.~Devine, Jon?, Cohen?, Others? \\
Sandia National Laboratories \\
Albuquerque, NM \\
sjplimp@sandia.gov
}

\date{}

\maketitle

\centerline{Keywords: MapReduce, message-passing, MPI, graph
algorithms, RMAT matrices}

\vspace*{0.4in}

\begin{abstract}

We describe a library that allows parallel algorithms to be expressed
in the MapReduce paradigm, where the user does not need to write
explicit parallel code, but instead provides a collection of ``map''
and ``reduce'' functions which operate on portions of a data set.  The
library itself uses MPI message passing calls to perform the needed
parallel operations across distributed memory.  We highlight how
MapReduce operations work in an MPI context, and also how they can be
coded in an out-of-core manner for data sets that do not fit within
the aggregate memory of a parallel machine.  Our motivation for
creating this library was to enable graph algorithms to be coded as
MapReduce operations, allowing processing of Terabyte-scale data sets.
We outline a MapReduce version of several such algorithms: vertex
ranking via PageRank, connected component identification, triangle
finding, Luby's algorithm for maximally independent sets, and
single-source shortest-path.  To input arbitrarily large artificial
graphs we generate RMAT matrices in parallel; we also provide a
MapReduce version of this operation.  Performance and scalability
results for the various algorithms are presented for graphs of varying
sizes on a distributed-memory cluster.  For some cases, we compare our
results with timings for different algorithms, different machines, and
different MapReduce software, namely Hadoop.  Our open-source library
is written in C++, is callable from C++, C, Fortran, or scripting
languages such as Python, and can run on any parallel platform that
supports MPI.

\end{abstract}

\pagebreak

\input{intro}
\input{mr}
\input{outcore}
\input{graph}
\input{results}
\input{thanks}

\bibliographystyle{aip}
\bibliography{paper}

\end{document}