### 2009-12-09 paper/thanks.tex

\section{Acknowledgements}

Sandia is a multiprogram laboratory operated by Sandia Corporation, a
Lockheed Martin Company, for the United States Department of
Energy's National Nuclear Security Administration under contract
DE-AC04-94AL85000.

Hadoop folks.