### 2010-03-26 paper/paper.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@478 3fc257bb-8550-0410-81b3-a5c103da83f7

\documentclass[11pt]{article}
\usepackage[pdftex]{graphicx}

\setlength{\textwidth}{6.5in}
\setlength{\textheight}{9.25in}

\setlength{\voffset}{-0.0in}
\hoffset=-50pt
\topmargin=0pt
\headheight=0pt
\headsep=0pt
\evensidemargin=72pt % 1in+7pt

\setlength{\parskip}{0.5ex}
\setlength{\parindent}{1em}
\setlength{\floatsep}{1pt}
\setlength{\textfloatsep}{2pt}
\setlength{\intextsep}{2pt}

\begin{document}

\title{Map{R}educe in {MPI} for Large-scale Graph Algorithms}

\author{
Steven J.~Plimpton, Karen D.~Devine, Jon?, Cohen?, Others? \\
Sandia National Laboratories \\
Albuquerque, NM \\
sjplimp@sandia.gov
}

\date{}

\maketitle

\centerline{Keywords: MapReduce, message-passing, MPI, graph
algorithms, RMAT matrices}

\vspace*{0.4in}

\begin{abstract}

We describe a parallel library that allows algorithms to be expressed
in the MapReduce paradigm, where the calling program does not need to
include explicit parallel code, but instead provides ``map'' and
``reduce'' functions which operate in parallel on elements of a data
set distributed across processors.  The library performs needed
parallel operations using message-passing (MPI) calls.  We describe
how typical MapReduce operations can be implemented in an MPI context,
and also in an out-of-core manner for data sets that do not fit within
the aggregate memory of a parallel machine.  Our motivation for
creating this library was to enable graph algorithms to be written as
MapReduce operations, allowing processing of Terabyte-scale data sets.
We outline MapReduce versions of several such algorithms: vertex
ranking via PageRank, triangle finding, connected component
identification, Luby's algorithm for maximally independent sets, and
single-source shortest-path.  To test the algorithms on arbitrarily
large artificial graphs we generate randomized RMAT matrices in
parallel; we also describe a MapReduce version of this operation.
Performance and scalability results for the various algorithms are
presented for graphs of varying sizes on a distributed-memory cluster.
For some cases, we compare the results with non-MapReduce algorithms,
different machines, and different MapReduce software, namely Hadoop.
Our open-source library is written in C++, is callable from C++, C,
Fortran, or scripting languages such as Python, and can run on any
parallel platform that supports MPI.

\end{abstract}

\pagebreak

\input{intro}
\input{mr}
\input{outcore}
\input{graph}
\input{results}
\input{thanks}

\bibliographystyle{aip}
\bibliography{paper}

\end{document}