### 2010-05-03 paper/intro.tex
Some edits; fixed explanations of additional MR-MPI functions; some of the function names were wrong.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@600 3fc257bb-8550-0410-81b3-a5c103da83f7

\section{Introduction}
\label{sec:intro}

MapReduce is the programming paradigm popularized by Google
researchers Dean and Ghemawat \cite{Dean}.  Their motivation was to
enable rapid development and deployment of analysis programs to
operate on massive data sets residing on Google's large distributed
clusters.  They introduced a novel way of thinking about certain kinds
of large-scale computations as a ``map'' operation followed by a
``reduce'' operation.  The power of the paradigm is that when cast in
this way, a nominally serial algorithm now becomes two highly parallel
operations working on data local to each processor, sandwiched around
an intermediate data-shuffling operation that requires inter-processor
communication.  The user need only write serial code for the
application-specific map and reduce functions; the parallel data
shuffle can be encapsulated in a library since its operation is
independent of the application.

The Google implementation of MapReduce is a C++ library with
communication between networked machines via remote procedure calls.
It allows for fault tolerance when large numbers of machines are used,
and can use disks as out-of-core memory to process petabyte-scale data
sets.  Tens of thousands of MapReduce programs have since been written
by Google researchers and are a significant part of the daily compute
tasks run by the company \cite{Dean2}.

Similarly, the open-source Hadoop implementation of MapReduce
\cite{Hadoop}, has become widely popular in the past few years for
parallel analysis of large-scale data sets at Yahoo and other
data-centric companies, as well as in university and laboratory
research groups, due to its free availability.  MapReduce programs in
Hadoop are typically written in Java, though it also supports use of
stand-alone map and reduce kernels, which can be written as shell
scripts or in other languages.

More recently, MapReduce formulations of traditional number-crunching
kinds of scientific computational tasks have been described, such as
post-processing analysis of simulation data \cite{Tu}, graph
algorithmics \cite{Cohen}, and linear algebra operations \cite{Fox}.
The paper by Tu et al \cite{Tu} was particularly insightful to us,
because it described how MapReduce could be implemented on top of the
ubiquitous distributed-memory message-passing interface (MPI), and how
the intermediate data-shuffle operation is conceptually identical to
the familiar MPI\_Alltoall operation.  Their implementation of
MapReduce was within a Python wrapper to simplify the writing of user
programs.  The paper motivated us to develop our own C++ library built
on top of MPI for use in graph analytics, which we initially released
as open-source software in mid-2009 \cite{MRMPI}.  We have since
worked to optimize several of the library's underlying algorithms and
to enable its operation in out-of-core mode on larger data sets.
These algorithmic improvements are described in this paper and are
part of the current downloadable version \cite{MRMPI}.

The MapReduce-MPI (MR-MPI) library described in this paper is a
simple, lightweight implementation of basic MapReduce functionality,
with the following features and limitations:

\begin{itemize}

\item {\it C++ library using MPI for inter-processor communication:}
The user writes a (typically) simple main program which runs on each
processor of a parallel machine, making calls to the MR-MPI library.
For map and reduce operations, the library calls back to user-provided
map() and reduce() functions.  The use of C++ allows precise control
over the memory and format of data allocated by each processor during
a MapReduce.  Library calls for performing a map, reduce, or data
shuffle, are synchronous, meaning all the processors participate and
finish the operation before proceeding.  Similarly, the use of MPI
within the library is the traditional mode of MPI\_Send and
MPI\_Recv calls between processor pairs using large aggregated
messages to improve bandwidth performance and reduce latency costs.  A
recent paper by \cite{Dongarra} also outlines the MapReduce formalism
from an MPI perspective, though they advocate a more asyncrhonous
approach, using one-way communication of small messages.

\item {Small, portable:} The entire MR-MPI library is a few thousand
lines of standard C++ code.  For parallel operation, the program is
linked with MPI, a standard message passing library available on all
distributed memory machines.  For serial operation, a dummy MPI
library (provided) can be substituted.  As a library, it can be
embedded in other codes \cite{Titan,TitanURL} to enable them to
perform MapReduces.

\item {\it In-core or out-of-core operation:} Each MapReduce object that a
processor defines allocates per-processor ``pages'' of memory, where
the page size is determined by the user.  Typical MapReduce operations
can be performed using a few such pages.  If the data set fits in a
single page (per processor), then the library performs its operations
in-core.  If the data set exceeds the page size, then processors each
write to temporary disk files (to local disk or a parallel file
system) as needed and subsequently read from them.  This allows
processing of data sets larger than the aggregate memory of all the
processors, i.e. up to the available aggregate disk space.

\item {\it Flexible programmability:} An advantage of writing a
MapReduce program on top of MPI, is that the user program can invoke
MPI calls directly, if desired.  For example, one-line calls to
MPI\_Allreduce are often useful in determining the status of an
iterative graph algorithm, as described in Section \ref{sec:graph}.
The library interface also provides a user data pointer as an argument
passed back to all callback functions, so it is easy for the user
program to store ``state'' on each processor, accessible during the
map and reduce operations.  For example, various flags can be stored
that alter the operation of a map or reduce function, as can richer
data structures, that accumulate the results.

\item {\it C++, C, and Python interfaces:} A C++ interface to the
MR-MPI library means a user program instantiates and then invokes
methods in one or more MapReduce objects.  A C interface means the
library can also be called from C or other high-level languages such as
Fortran.  A C interface also means the library can be easily wrapped
by Python via the Python ``ctypes'' module.  The library can then be
called from a Python script, allowing the user to write map() and
reduce() callback functions in Python.  If a machine supports running
Python in parallel, a parallel MapReduce can also be run in this mode.

\item {No fault tolerance:} Current MPI implementations do not enable
easy detection of a dead processor, or retrieval of the data it was
working on.  So like most MPI programs, a parallel program calling the
MR-MPI library will hang or crash if a processor goes away.  Unlike
Hadoop, and its HDFS file system which provides for data redundancy,
the MR-MPI library simply reads and writes simple, flat files.  It can
use local per-processor disks, or a parallel file system, if
available, but these typically provide no data redundancy.

\end{itemize}

The remainder of the paper is organized as follows.  The next two
sections \ref{sec:mr} and \ref{sec:outcore} describe how in-core and
out-of-core MapReduce primitives are formulated as MPI-based
operations in the MR-MPI library.  Section \ref{sec:graph} briefly
describes the formulation of several common graph algorithms as
MapReduce operations.  Section \ref{sec:results} gives performance
results for these algorithms running on a parallel cluster for graphs
%KDDKDD CHECK SIZES
ranging in size from 1 million to 1 trillion vertices or edges.  In
this section, we highlight the performance and complexity trade-offs
of a MapReduce approach versus other more special-purpose algorithms.
The latter generally perform better but are harder to implement
efficiently on distributed memory machines, due to the required
explicit management of parallelism, particularly for large out-of-core
data sets.  Section~\ref{sec:lessons} summarizes some lessons learned
from the implementation and use of our library.