### 2010-05-26 paper/thanks.tex

\section{Acknowledgements}
\label{sec:thanks}

We thank the following individuals for their contributions to this
paper: Greg Bayer and Todd Plantenga (Sandia) for explaining Hadoop
concepts to us, and for the Hadoop implementations and timings of
Section~\ref{sec:results}; Jon Cohen (NSA) for fruitful discussions
about his MapReduce graph algorithms~\cite{Cohen09}; Brian Barrett
(Sandia) for the PBGL results of Section~\ref{sec:results}; Jon Berry
(Sandia) for the MTGL results of Section~\ref{sec:results}, and for
his overall support of this work and many useful discussions.

Sandia National Laboratories is a multi-program laboratory operated by
Sandia Corporation, a wholly owned subsidiary of Lockheed Martin
company, for the U.S. Department of Energy's National Nuclear Security
Administration under contract DE-AC04-94AL85000.