### 2010-06-07 paper/lessons.tex

\section{Lessons Learned}
\label{sec:lessons}

We conclude with several observations about performing MapReduce
operations on distributed-memory parallel machines via MPI.

MapReduce achieves parallelism through randomizing the distribution of
data across processors, which often intentionally ignores data
locality.  This translates into maximal data movement (during a
shuffle) with communication between all pairs of processors.  But the
benefit is often good load-balance, even for hard-to-balance irregular
data sets.  By contrast, more traditional distributed-memory parallel
algorithms, e.g. for matrix operations, or grid-based or
particle-based simulation codes, tend to work hard to localize data
and minimize communication.  To do this they typically require a lot
of application-specific logic and parallel communication coding to
create and maintain a data decomposition, generate ghost versions of
nearby spatially-decomposed data, etc.

MapReduce algorithms can be hard to design, but are often relatively
easy to write and debug.  Thinking about a computational task from a
MapReduce perspective is different from traditional distributed-memory
parallel algorithm design.  For example, with an MPI mindset, it often
seems heretical to intentionally ignore data locality.  However,
writing small {\it map()} and {\it reduce()} functions is typically easy.  And
writing an algorithm that involves complex parallel operations
without actually needing to write application-specific parallel code
to move data via MPI calls is often a pleasant surprise.  Moreover,
if the MapReduce algorithm is initially coded so that it runs
correctly on one processor, it often works out-of-the-box on hundreds
or thousands of processors, without the need for additional debugging.

Performing MapReduce operations on a fixed allocation of processors on
a traditional MPI-based parallel machine is a somewhat different
conceptual model than that of cloud-computing MapReduces using (for
example) Hadoop.  In the former case, one can potentially control
which processor owns which data at various stages of an algorithm.
This is somewhat hidden from the user in a typical cloud-computing
model, where data simply exists somewhere in the cloud and Hadoop
ensures data moves where it is needed and is operated on by some
processor.  The cloud model is a nice data-centric abstraction which
allows for fault tolerance both to data loss (via redundant storage)
and to processor failure (via reassignment of work), neither of which
is typically possible on current MPI-based parallel machines.

However, since the MPI implementation of MapReduce as
described in this paper is processor-centric, one can sometimes
fruitfully exploit the possibility for processors to maintain
``state'' over the course of multiple map and reduce operations.  By
controlling where data resides for maps and reduces (e.g. via a
user-specified hash function), and by assuming that processor will
always be available, more efficient operations with less data movement
are sometimes possible.  The discussion of enhanced graph algorithms
in Section \ref{sec:graph} illustrated this.  To fully exploit this
idea for large data sets, mechanisms are needed for processors to
store, retrieve, and efficiently find needed datums on local disk
(e.g. static edges of a graph), so that archived state can be used
efficiently during subsequent map and reduce operations.

Finally, though this paper focuses on graph algorithms expressed as
MapReduce operations, there is nothing about the MR-MPI library itself
that is graph centric.  We hope the library can be generally useful on
large-scale monolithic or cloud-style parallel machines that support
MPI, for a variety of data-intense or compute-intense problems that
are amenable to solution using a MapReduce paradigm.