### 2010-06-07 paper/lessons.tex

\section{Lessons Learned}
\label{sec:lessons}

We conclude with several observations about performing MapReduce
operations on distributed-memory parallel machines via MPI.

MapReduce achieves parallelism through randomizing the distribution of
data across processors, which often intentionally ignores data
locality.  This translates into maximal data movement (during a data
shuffle or {\it collate()} operation) with communication between all
pairs of processors.  But the benefit is often good load-balance, even
for hard-to-balance irregular data sets.  By contrast, more
traditional distributed-memory parallel algorithms, e.g. for matrix
operations, or grid-based or particle-based simulation codes,
typically work hard to localize data and minimize communication.  To
do this they often require a lot of application-specific logic and
parallel communication coding to create and maintain a data
decomposition, generate ghost versions of nearby spatially-decomposed
data, etc.

MapReduce algorithms can be hard to design, but are often relatively
easy to write and debug.  Thinking about a computational task from a
MapReduce perspective is different from traditional distributed-memory
parallel algorithm design.  For example, with an MPI mindset, it often
seems heretical to intentionally ignore data locality.  However,
writing small {\it map()} and {\it reduce()} functions is typically
easy.  And writing an algorithm that involves complex parallel
operations without needing to write application-specific code to
communicate data via MPI calls or to move data between out-of-core
disk files and processor memory is often a pleasant surprise.
Moreover, if the MapReduce algorithm is initially coded so that it
runs correctly on one processor, it often works out-of-the-box on
hundreds or thousands of processors, without the need for additional
debugging.

Performing MapReduce operations on a fixed allocation of processors on
a traditional MPI-based parallel machine is a somewhat different
conceptual model than that of performing MapReduce operations on a
cloud computer using (for example) Hadoop.  In the former case, one
can potentially control which processor owns which data at various
stages of an algorithm.  This is somewhat hidden from the user in a
typical cloud-computing model, where data simply exists somewhere in
the cloud and Hadoop ensures data moves where it is needed and is
operated on by some processor.  The cloud model is a nice data-centric
abstraction which allows for fault tolerance both to data loss (via
redundant storage) and to processor failure (via reassignment of
work), neither of which is typically possible on current MPI-based
parallel machines.

However, since the MPI implementation of MapReduce as described in
this paper is processor-centric, one can sometimes fruitfully exploit
the possibility for processors to maintain ``state'' over the course
of multiple map and reduce operations.  By controlling where data
resides for maps and reduces (e.g. via a user-specified hash
function), and by assuming that processor will always be available,
more efficient operations with less data movement are sometimes
possible.  The discussion of enhanced graph algorithms in Section
\ref{sec:graph} illustrated this.  To fully exploit this idea for
large data sets, mechanisms and data structures are needed for
processors to store, retrieve, and efficiently find needed datums on
local disk (e.g. static edges of a graph), so that archived state can
be used efficiently during subsequent map and reduce operations.

Finally, though this paper focuses on graph algorithms expressed as
MapReduce operations, there is nothing about the MR-MPI library itself
that is graph specific.  We hope the library can be generally useful
on large-scale monolithic or cloud-style parallel machines that
support MPI, for a variety of data-intensive or compute-intensive
problems that are amenable to solution using a MapReduce paradigm.