### 2010-06-07 paper/results.tex

\section{Performance Results}
\label{sec:results}

In this section, we present performance results for the MapReduce
graph algorithms of the preceeding section, implemented as small C++
programs calling our MR-MPI library.  The benchmarks were run on a
medium-sized Linux cluster of 2 GHz dual-core AMD Opteron processors
connected via a Myrinet network.  Most importantly for MR-MPI, each
node of the cluster has one local disk, which is used for out-of-core
operations in MR-MPI.  To avoid contention for disk I/O, we ran all
experiments with one MPI process per node.  For comparisons with other
implementations, we used either the same cluster or, where noted,
Sandia's Cray XMT, a multi-threaded parallel computer with 500 {MHz}
processors and a 3D-Torus network.

We ran each of the algorithms on three R-MAT graphs of different
sizes, each on a varying number of processors.  Details of the input
data are shown in Table~\ref{t:rmats}.  The {\it small} problem
(around 8M edges) can typically be run on a single processor without
incurring out-of-core operations.  The {\it medium} problem (around
134M edges) can be run on a single processor with out-of-core
operations; larger processor configurations, however, do not
necessarily require out-of-core operations.  The {\it large} problem
(around 2B edges) requires out-of-core operations on our 64-node
cluster.

%The {\it x-large} problem 
%(around 34B edges) requires most of the machine to run.

All data sets used R-MAT parameters $(a, b, c, d) = (0.57, 0.19, 0.19,
0.05)$ and generated 8 edges per vertex (on average).  These values
create a highly skewed degree distribution, as indicated by the
maximumn vertex degree in Table~\ref{t:rmats}.

\begin{table}
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|c|c|}
\hline
Data & \# of    & \# of & Maximum \\
Set  & vertices & edges & vertex degree\\
\hline
RMAT-20 (small)   &$2^{20} \approx 1M$ & $2^{23} \approx 8M$ &  $\approx 24K$ \\
RMAT-24 (medium)  &$2^{24} \approx 17M$ & $2^{27} \approx 134M$ &  $\approx 147K$ \\
RMAT-28 (large)   &$2^{28} \approx 268M$ & $2^{31} \approx 2B$& $\approx 880K$ \\
%RMAT-32 (x-large) &$2^{32} \approx 4B$ & $2^{35} \approx 34B$ &   \\
\hline
\end{tabular}
\caption{Characteristics of R-MAT input data for graph algorithm
benchmarks.}
\label{t:rmats}
\end{center}
\end{table}

The resulting timings give a sense of the inherent scalability of the
MapReduce algorithms as graph size grows on a fixed number of
processors, and of the parallel scalability for computing on a graph
of fixed size on a growing number of processors.  Where available, we
compare the MapReduce algorithm with other parallel implementations,
including more traditional distributed-memory algorithms and
multi-threaded algorithms in the Multi-Threaded Graph Library
(MTGL)~\cite{MTGL} on the Cray XMT.  We compute parallel efficiency on
$P$ processors as $({time}_{M} \times M) / ({time}_P \times P)$ where
$M$ is the smallest number of processors on which the experiment was
run, and ${time}_I$ is the execution time required on $I$ processors.

\subsection{R-MAT generation results}

In Figure~\ref{f:rmat}, we show the scalability of the R-MAT
generation algorithm (Figure~\ref{fig:rmat2}) for RMAT-20, RMAT-24 and
RMAT-28.  For RMAT-20 and RMAT-24, superlinear speed-up is shown.
This speed-up is due to the decreased amount of file I/O needed with
greater numbers of processors; with a larger total memory, a
fixed-size problem requires less I/O with more processors.  For
RMAT-28, which requires significant out-of-core operations, parallel
efficiency ranges from 52\% to 97\%.  Comparisons between the enhanced
algorithm (Figure~\ref{fig:rmat2}) and the original algorithm
(Figure~\ref{fig:rmat}) showed approximately 10\% reduction in
execution time for the enhanced algorithm.

%Also in Figure~\ref{f:rmat}, we compare our MR-MPI implementation with an 
%R-MAT generator in MTGL on the Cray XMT.  In the MTGL implementation, 
%threads generate edges independently.  A global hash table is used to remove
%duplicate edges.  Our MR-MPI implementation runs more quickly than the MTGL
%implementation.  The differences in execution time are greatest on 64
%processors, where MR-MPI needs the least file I/O for out-of-core operations.

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_rmat.pdf}
\caption{Performance of the MR-MPI R-MAT generation algorithm (Figure~\ref{fig:rmat2}) 
%and an MTGL implementation.
}
\label{f:rmat}
\end{figure}

\subsection{PageRank results}
\label{subsec:results_pagerank}

In Figure~\ref{f:pr}, we show the performance of the MR-MPI PageRank
algorithm (Figure~\ref{fig:pr2}) compared to a distributed-memory
matrix-based implementation using the linear algebra toolkit
Trilinos~\cite{Trilinos-Overview}.  The matrix-based
distributed-memory implementation of PageRank uses Trilinos Epetra
matrix/vector classes to represent the graph and PageRank vector.
Rows of matrix $A$ and the associated entries of the PageRank vector
$x$ are uniquely assigned to processors; a random permutation of the
input matrix effectively load balances the non-zero matrix entries
across processors.  Interprocessor communication gathers $x$ values
for matrix-vector multiplication and sums partial products into the
$y$ vector.  Most communication is point-to-point communication, but
some global communication is needed for computing residuals and norms
of $x$ and $y$.

Figure~\ref{f:pr} shows the execution time per PageRank iteration for
R-MAT matrices RMAT-20, RMAT-24 and RMAT-28.  Converging the PageRank
iterations to tolerance $0.002$ requires five or six iterations.
Several R-MAT matrices of each size were generated for the
experiments; the average time over the matrices is reported here.  The
Trilinos implementations show near-perfect strong scaling for RMAT-20
and RMAT-24.  The MR-MPI implementations also demonstrate good strong
scaling.  However, MR-MPI's execution time is at least an order of
magnitude greater than the Trilinos implementation.  This result is
due to two factors: ({\it i}) a higher volume of communication in the
MapReduce implementation (where all edges are communicated in each
iteration), compared to the Trilinos implementation (where only
vertex-based data is communicated), and ({\it ii}) out-of-core
operations in MR-MPI were performed because of the page-size
restriction, while all Trilinos operations were performed in-core.
The benefit of MR-MPI's out-of-core implementation is seen, however,
with the RMAT-28 data set, which could be solved on smaller processor
sets than the Trilinos implementation.  For these experiments, the
Trilinos implementation required 64 processors for the RMAT-28 data
set, since it will not operate on matrices that exceed the aggregate
memory of the processors.

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_pagerank.pdf}
\caption{Comparison of PageRank implementations using 
MR-MPI (Figure~\ref{fig:pr2}) and 
Trilinos' matrix/vector classes on R-MAT data sets.}
\label{f:pr}
\end{figure}

\subsection{Triangle finding results}

In Figure~\ref{f:tri}, we show the performance of the triangle finding
algorithm (Figure~\ref{fig:tri}).  Execution times for this algorithm
were too large to allow the problem to be easily run with RMAT-28 on
our cluster.  Parallel efficiencies for RMAT-20 ranged from 80\% to
140\%; for RMAT-24, they ranged from 50\% to 120\%.

%Comparisons with an MTGL implementation on the 
%Cray XMT give mixed results.  When MR-MPI uses the least file I/O for 
%out-of-core operations (e.g., RMAT-20 on 32 and 64 processors), it 
%outperforms the MTGL implementation.  However, when significant out-of-core
%operations are needed (e.g., for RMAT-24), the MTGL implementation is faster.

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_tri.pdf}
\caption{Performance of the MR-MPI triangle-finding algorithm
(Figure~\ref{fig:tri}).}
\label{f:tri}
\end{figure}

\subsection{Connected Components}

In Figure~\ref{fig:cc}, we show the performance of the connected
component identification algorithm (Figure~\ref{fig:cc}).  A
comparison is made with a hybrid ``Giant Connected Component''
implementation using both Trilinos and MR-MPI.  Power law graphs often
have one or two very large components and many very small components.
The hybrid algorithm exploits this feature of the data by using
inexpensive breadth-first search (BFS) from the vertex with highest
degree to identify the largest components, followed by a more
expensive algorithm to identify the small components in the remainder
of the graph.  In our hybrid implementation, we intially perform
matrix-vector multiplications in Trilinos to perform a BFS, finding
components that include (in total) 70\% or more of the vertices.  We
then apply our MR-MPI algorithm to the remaining graph to identify the
small components.  This hybrid approach is quite effective, reducing
the execution time to identify all components of RMAT-20 and RMAT-24
by 80-99\%, as shown in Figure~\ref{f:cc}.  The benefit of MR-MPI is
seen, however, for RMAT-28, where the graph is too large to fit into
memory, and thus Trilinos cannot perform the initial BFS for our
hybrid algorithm.

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_cc.pdf}
\caption{Performance of the MR-MPI connected components algorithm (Figure~\ref{fig:cc}) compared with a hybrid ``Giant Connected Component'' algorithm based
on Trilinos.}
\label{f:cc}
\end{figure}

\subsection{Maximally independent set results}

The execution times for the maximal independent set algorithm
(Figure~\ref{fig:luby}) are shown in Figure~\ref{f:luby}.  Like the
R-MAT generation results, superlinear speed-up of the algorithm occurs
for RMAT-20 and RMAT-24, as more of the graph fits into processor
memory and less file I/O is needed.  For RMAT-28, the algorithm
requires significant out-of-core operations. In this case, parallel
efficiency is nearly perfect going from 8 to 64 processors.

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_luby.pdf}
\caption{Performance of the MR-MPI maximal independent set algorithm (Figure~\ref{fig:luby}).}
\label{f:luby}
\end{figure}

\subsection{Single-source shortest path results}

The execution times for the single-source shortest path algorithms
(Figures~\ref{fig:sssp} and \ref{fig:sssp2}) are shown in
Figure~\ref{f:sssp}.  The results show the benefit of using the
enhanced algorithm, providing at least a 17\% (and often greater)
reduction in execution time due to reduced communication; only the
updated distances are communicated throughout most of the enhanced
algorithm.  However, the execution times are still large compared to
multi-threaded implementations; for example, Madduri et
al.~\cite{Madduri07} report execution times of only 11 seconds on 40
processors for a multithreaded implementation on a Cray MTA-2 (the
predecessor of the XMT) on graphs with one billion edges.

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_sssp.pdf}
\caption{Execution times for SSSP using MR-MPI with R-MAT matrices.  
Both the original algorithm (Figure~\ref{fig:sssp}) and the enhanced 
algorithm (Figure~\ref{fig:sssp2}) are included.}
\label{f:sssp}
\end{figure}

To compare our MR-MPI implementation with a wider set of algorithms,
we performed experiments comparing MR-MPI, Hadoop, PBGL (Parallel
Boost Graph Library)~\cite{PBGL} and a multi-threaded implementation
on the Cray XMT using two web graphs: {WebGraphA} with 13.2M vertices
and 31.9M edges, and {WebGraphB} with 187.6M vertices and 531.9M
edges.  In Figure~\ref{f:ssspA}, we show execution times for the SSSP
algorithm using {WebGraphA}.  Like our MR-MPI implementation, the
Hadoop implementation is a Bellman-Ford-style~\cite{Bellman58,Ford62}
algorithm.  The XMT and PBGL implementations are based on
delta-stepping~\cite{MeyerSanders98}, and do not require full
iterations over the entire edge list to advance a breadth-first
search.  We observe that the MR-MPI implementation runs in less time
than the Hadoop implementation, but requires significantly more time
than the XMT and PBGL implementations.  In experiments with
{WebGraphB}, the benefit of the enhanced algorithm
(Figure~\ref{fig:sssp2}) is clearly shown, with a 40\% reduction in
execution time compared to the original SSSP algorithm
(Figure~\ref{fig:sssp}).  But the Bellman-Ford-style iterations are
especially harmful to the MR-MPI and Hadoop implementations for
{WebGraphB}, which required 110 iterations to complete; execution
times for this data set are shown in Table~\ref{t:ssspB}.

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_ssspA.pdf}
\caption{SSSP using MR-MPI for WebGraphA with
13.2M vertices and 31.9M edges.  Runtimes using Hadoop and PBGL
are also shown.}
\label{f:ssspA}
\end{figure}

\begin{table}
\begin{center}
\begin{tabular}{|l|c|r|}
\hline
Implementation & Number of Processes & SSSP Execution Time \\
\hline
Hadoop & 48  & 38,925 secs.\\
MR-MPI original (Figure~\ref{fig:sssp}) & 48 &  13,505 secs.\\
MR-MPI enhanced (Figure~\ref{fig:sssp2}) & 48 &  8,031 secs.\\
XMT/C  & 32 &  37 secs.\\
%Out-of-core MR-MPI & 96 &   8,358 secs.\\
%Out-of-core MR-MPI & 64 &  12,882 secs.\\
%Out-of-core MR-MPI & 100 &  6,280 secs.\\
\hline
\end{tabular}
\caption{Execution times for SSSP with {WebGraphB}.}
\label{t:ssspB}
\end{center}
\end{table}

\subsection{Scalability to large numbers of processors}

Finally, we demonstrate the scalability of our MR-MPI library to large
numbers of processors.  The library was used on Sandia's Redstorm and
Thunderbird parallel computers.  Redstorm is a large Cray XT3 with 2+
{GHz} dual/quad-core AMD Opteron processors and a custom interconnect
providing 9.6 GB/s of interprocessor bandwidth.  Thunderbird is a
large Linux cluster with 3.6 {GHz} dual-core Intel EM64T processors
connected by an Infiniband network.  Because these systems do not have
local disks for each processor, we selected a data set and page sizes
that fit in memory, so out-of-core operations were not needed.  For
these experiments, we used an R-MAT data set with with $2^{25}$
vertices and $2^{28}$ edges, with parameters given in
Table~\ref{t:rmat}.  We ran both the PageRank and Connected Components
algorithms.

\begin{table}
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|c|c|}
\hline
Data & R-MAT  & R-MAT  & R-MAT  & R-MAT  & \# of    & \# of & Maximum \\
Set  & a      & b      & c      & d      & vertices & edges & vertex degree\\
\hline
nice  & 0.45 & 0.15 & 0.15 & 0.25 & $2^{25}$ & $2^{28}$ & 1108 \\
nasty & 0.57 & 0.19 & 0.19 & 0.05 & $2^{25}$ & $2^{28}$ & 230,207\\
\hline
\end{tabular}
\caption{Characteristics of R-MAT input data for PageRank and Connected
Components scalability experiments.}
\label{t:rmat}
\end{center}
\end{table}

Figure~\ref{f:prbig}, shows the performance of the various PageRank
implementations on distributed memory and multi-threaded
architectures.  The MR-MPI and Trilinos implementations are described
in Sections~\ref{subsec:graph_pagerank}
and~\ref{subsec:results_pagerank}, respectively.  In the
multi-threaded MTGL implementation, rank propagates via adjacency list
traversal in a compressed sparse-row data structure.  To maintain its
scalability, code must be written so that a single thread spawns the
loop that processes all in-neighbors of a given vertex; this detail
enables the compiler to generate hotspot-free code.

The MR-MPI implementation demonstrated good scalability up to 1024
processors; however, as before, it required an order-of-magnitude more
execution time than the matrix-based implementations on Redstorm.  The
distributed memory matrix-based implementations are competitive with
the multi-threaded implementation in MTGL on the Cray XMT.

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_pagerank_big.pdf}
\caption{Scalability comparison of PageRank using MapReduce (MR-MPI),
matrix-based (Trilinos), and multi-threaded (MTGL) implementation on
the R-MAT data sets in Table~\ref{t:rmat}.}
\label{f:prbig}
\end{figure}

Similar results were obtained for the Connected Components algorithm,
as shown in Figure~\ref{f:ccbig}.  As with PageRank, the MR-MPI
implementation showed good scalability up to 1024 processors, but
required significantly more time than the hybrid algorithm using
Trilinos and MR-MPI or the MTGL algorithm.

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_cc_big.pdf}
\caption{Scalability comparison of Connected Components algorithms using 
MapReduce (MR-MPI),
matrix-based/MapReduce hybrid (Trilinos/MR-MPI), and MTGL implementations
on the R-MAT data sets in Table~\ref{t:rmat}.}
\label{f:ccbig}
\end{figure}