### 2010-11-03 paper/mr.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@659 3fc257bb-8550-0410-81b3-a5c103da83f7

\section{MapReduce in MPI}
\label{sec:mr}

The basic datums stored and operated on by any MapReduce framweork are
key/value (KV) pairs.  In the MR-MPI library, individual keys or
values can be of any data type or length, or combinations of multiple
types (one integer, a string of characters, two integers and a double,
etc); they are simply treated as byte strings by the library.  A KV
pair always has a key; its value may be NULL.  KV pairs are stored
within a MapReduce (MR) object.  A user program may create one or more
MR objects to implement an algorithm.  Various MapReduce operations
(map, reduce, etc) are invoked on an object and its KV pairs; KV pairs
can also be passed between and combined across objects.

A related data type in the MR-MPI library is the key/multivalue (KMV)
pair, where all values associated with the same key are collected and
stored contiguously as a multivalue, which is just a longer byte
string with an associated vector of lengths, one integer length per
value.  In this section, we assume the KV and KVM datums operated on
by each processor fit in local memory.

A typical MR-MPI program makes at least three calls to the MR-MPI
library, to perform {\it map()}, {\it collate()}, and {\it reduce()}
operations on a MR object it creates.  In a {\it map()}, zero or more
key/value pairs are generated by each processor.  Often this is done
using data read from files, but a {\it map()} may generate data itself
or process existing KV pairs to create new ones.  The KV pairs
produced are stored locally by each processor; a {\it map()} thus
requires no inter-processor communication.  Users call the library
with a count of tasks to perform and a pointer to a user function; the
MR-MPI {\it map()} operation invokes the user function multiple times
as a callback.  Depending on which variant of {\it map()} is called,
the user function may be passed a file name, a chunk of bytes from a
large file, a task ID, or a KV pair.  Options for assigning map tasks
to processors are specified by the user and include assigning
consecutive chunks of tasks to processors, striding the tasks across
processors, or using a master-slave model that is useful when tasks
have widely varying workloads.

The {\it collate()} operation (or data shuffle in Hadoop) identifies
unique keys and collects all the values associated with those keys to
create KMV pairs.  This is done in two stages, the first of which
requires communication, since KV pairs with the same key may be owned
by any processor.  Each processor hashes each of its keys to determine
which processor will ``own'' it.  The $k$-byte length key is hashed
into a 32-bit value whose remainder modulo $P$ processors is the
owning processor rank.  Alternatively, the user can provide a hash
function which converts the key into a processor rank.  Each processor
then sends each of its KV pairs to the owning processor.

After receiving new KV pairs, the second stage is an on-processor
computation, requiring no further communication.  Each processor
reorganizes its KV pairs into KMV pairs, one for each unique key it
owns.  This is done using a hash table, rather than a sort.  A sort
scales as $N\log_2(N)$, i.e. it requires $\log_2(N)$ passes through
the $N$ KV pairs.  With hashing, the list of KMV pairs can be created
in two passes.  The first pass populates the hash table with needed
count and length information; the second pass copies the key and value
datums into the appropriate location in a new KMV data structure.
Since the cost to lookup a key in a well-formed hash table is a
constant-time $O(1)$ operation, the cost of the data reorganization is
also $O(N)$.  This methodology has the added benefit that the number
of values in each KMV pair is known and can be passed to the user
function during a reduce.  For some reduce operations, this count is
all the information a reduce requires; the values need not be looped
over.  The count is not available in Hadoop-style data shuffles, which
sort the values; counting the values associated with the same key
requires iterating over the values.

Note that the first portion of the {\it collate()} operation involves
all-to-all communication since each processor exchanges KV pairs with
every other processor.  The communication can either be done via a
MPI\_Alltoall() library call, or by a custom routine that aggregates
datums into large messages and invokes point-to-point MPI\_Send() and
MPI\_IRecv() calls.

The {\it reduce()} operation processes KMV pairs and can produce new
KV pairs for continued computation.  Each processor operates only on
the KMV pairs it owns; no communication is required.  As with the {\it
map()}, users call the library with a pointer to a user function.  The
MR-MPI {\it reduce()} operation invokes the user function, once for
each KMV pair.

Several related MapReduce operations are provided by the library.  For
example, the {\it collate()} function described above calls two other
functions: {\it aggregate()}, which performs the all-to-all
communication, and {\it convert()}, which reorganizes a set of KV
pairs into KMV pairs.  Both functions can be called directly.  The
{\it compress()} function allows on-processor operations to be
performed; it is equivalent to a {\it convert()} with on-processor KVs
as input, followed by a {\it reduce()}.  The {\it clone()} function
turns a list of KV pairs into KMV pairs, with one value per key.  The
{\it collapse()} function turns $N$ KV pairs into one KMV pair, with
the keys and values of the KV pairs becoming $2N$ values of a single
multivalue assigned to a new key.  The {\it gather()} function
collects KV pairs from all processors to a subset of processors; it is
useful for doing output from one or a few processors.  Library calls
for sorting datums by key or value or for sorting the values within
each multivalue are also provided.  These routines invoke the
C-library {\it quicksort()} function to compute the sorted ordering of
the KV pairs (or values within a multivalue), using a user-provided
comparison function.  The KV pairs (or values in a multivalue) are
then copied into a new data structure in sorted order.

The interface to various low-level operations is provided so that a
user's program can string them together in various ways to construct
usefule MapReduce algorithms.  For example, output from a {\it
reduce()} can serve as input to a subsequent {\it map()} or {\it
collate()}.  Likewise, data can be partitioned across multiple
MapReduce (MR) objects, each with their own set of KV pairs.  For
example, as will be illustrated in Section ~\ref{sec:graph}, one MR
object can store the edges of a graph, and another its vertices.  KV
pairs from multiple MR objects can be combined to perform new
sequences of {\it map()}, {\it collate()}, and {\it reduce()}
operations.

The above discussion assumed that the KV or KMV pairs stored by a
processor fit in its physical memory.  In this case, the MR-MPI
library performs only ``in-core'' processing and no disk files are
written or read by any of the processors, aside from initial input
data (if it exists) or final output data (if it is generated).  Note
that the aggregate physical memory of large parallel machines can be
multiple terabytes, which allows for large data sets to be processed
in-core, assuming the KV and KMV pairs remain evenly distributed
across processors throughout the sequence of MapReduce operations.
The use of hashing to assign keys to processors typically provides for
such load-balancing.  In the next section, we discuss what happens
when data sets do not fit in available memory.