### 2010-11-04 paper/graph.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@662 3fc257bb-8550-0410-81b3-a5c103da83f7

\section{Graph Algorithms in MapReduce}
\label{sec:graph}

We begin with a MapReduce procedure for creating large, sparse,
randomized graphs, since they are the input for the algorithms
discussed below.  R-MAT matrices ~\cite{RMAT} are recursively
generated graphs with power-law degree distributions.  They are
commonly used to represent web and social networks.  The user
specifices six parameters that define the graph: the number of
vertices $N$ and edges $M$, and four parameters $a$, $b$, $c$, $d$
that sum to 1.0 and are discussed below.  The algorithm in
Figure~\ref{fig:rmat} uses a single MapReduce object to generate $M$
unique non-zero entries in a sparse $N \times N$ matrix $G$, where
each entry $G_{ij}$ represents an edge between graph vertices
$(V_i,V_j)$, with $V_i$ an integer from $1$ to $N$.

\begin{figure}[htb]
 \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
  xxxx\=xxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill

Input: \\
\> MapReduce object $G$ = edges of graph, initially empty \\
\\
$M_{\rm remain} = M$ \\
while $M_{\rm remain} > 0$: \\
\> Map $G$: \> Generate $M_{\rm remain}/P$ random edges $(V_i,V_j)$ on each processor \\
\> \> Output: Key = $(V_i,V_j)$, Value = NULL \\
\> Collate $G$ \\
\> Reduce $G$: \> Remove duplicate edges \\
\> \> Input: Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
\> \> Output: Key = $(V_i,V_j)$, Value = NULL \\
\> $M_{\rm remain} = M - N_{kv}$ \\
\\
Output: \\
\> MapReduce object $G$ = edges of graph: Key = $(V_i,V_j)$, Value = NULL \\

  \end{tabbing}
 \end{minipage}}\end{center}

 \caption{\it MapReduce algorithm for R-MAT graph generation on $P$
 processors, using a single MapReduce object $G$.}

 \label{fig:rmat}
\end{figure}

In the {\it map()} operation, each of $P$ processors generates a $1/P$
fraction of the desired edges.  A single random edge $(V_i,V_j)$ is
computed recursively as follows.  Pick a random quadrant of the $G$
matrix with relative probabilities $a$, $b$, $c$, and $d$.  Treat the
chosen quadrant as a sub-matrix and select a random quadrant within
it, in the same manner.  Repeat this process $n$ times where $N =
2^n$.  At the end of the recursion, the final ``quadrant'' is a single
non-zero matrix element $G_{ij}$.

Though $G$ is very sparse, the {\it map()} typically generates some
small number of duplicate edges.  The {\it collate()} and {\it
reduce()} operations remove the duplicates.  The entire
map-collate-reduce sequence is repeated until the number of resulting
key/value (KV) pairs $N_{kv} = M$.  For reasonably sparse graphs this
typically takes only a few iterations.

Note that the degree distribution of vertices in the graph depends on
the choice of parameters $a$, $b$, $c$, $d$.  If one of the four
values is larger than the other three, a skewed distribution results.
Variants of the above algorithm can be used when $N$ is not a
power-of-two, to generate graphs with weighted edges (assign a numeric
value to the $G_{ij}$ edge), graphs without self edges (require $i \ne
j$), or graphs with undirected edges (require $i < j$).  Or the
general R-MAT matrix can be further processed by MapReduce operations
to meet these requirements.  For some algorithms below (e.g., triangle
finding and maximal independent set generation), we post-process R-MAT
matrices to create upper-triangular R-MAT matrices representing
undirected graphs.  For each edge $(V_i,V_j)$ with $i > j$, we add
edge $(V_j,V_i)$, essentially symmetrizing the matrix; we then remove
duplicate edges and self edges, leaving only edges with $i < j$.

By using an additional MapReduce object, we can improve the
performance of the R-MAT generation algorithm, particularly when
out-of-core operations are needed.  In the enhanced algorithm
(Figure~\ref{fig:rmat2}), we emit newly created edges into MapReduce
object $G_{new}$ and {\it aggregate()} the edges to processors.  We
then add those edges to the MapReduce object $G$, containing all
previously generated unique edges; the edges of $G$ are already
aggregated to processors using the same mapping as $G_{new}$.  We can
then use the {\it compress()} operation to reduce duplicates locally
on each processor.  The cost of the enhanced algorithm is less due to
the smaller number of KV pairs communicated in the {\it aggregate()},
compared to the {\it collate()} operation of the original algorithm,
particularly in later iterations of the algorithm where few new edges
are generated.

\begin{figure}[htb]
 \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
  xxxx\=xxxxxxxxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill

Input: \\
\> MapReduce object $G$ = edges of graph, initially empty \\
\> MapReduce object $G_{new}$ = new edges of graph, initially empty \\
\\
$M_{\rm remain} = M$ \\
while $M_{\rm remain} > 0$: \\
\> Map $G_{new}$: \> Generate $M_{\rm remain}/P$ random edges $(V_i,V_j)$ on each processor \\
\> \> Output: Key = $(V_i,V_j)$, Value = NULL \\
\> Aggregate $G_{new}$ \\
\> Add $G_{new}$ to $G$, empty $G_{new}$ \\
\> Compress $G$: Remove duplicate edges \\
\> \> Input: Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
\> \> Output: Key = $(V_i,V_j)$, Value = NULL \\
\> $M_{\rm remain} = M - N_{kv}$ \\
\\
Output: \\
\> MapReduce object $G$ = edges of graph: Key = $(V_i,V_j)$, Value = NULL \\

  \end{tabbing}
 \end{minipage}}\end{center}

 \caption{\it Enhanced MapReduce algorithm for R-MAT graph generation
 on $P$ processors, using two MapReduce objects $G$ and $G_{new}$.}

 \label{fig:rmat2}
\end{figure}

\subsection{PageRank}
\label{subsec:graph_pagerank}

The PageRank algorithm assigns a relative numeric rank to each vertex
in a graph.  In an Internet context, it models the web as a directed
graph $G(V,E)$, with each vertex $V_i \in V$ representing a web page
and each edge $(V_i, V_j) \in E$ representing a hyperlink from $V_i$
to $V_j$.  The probability of moving from $V_i$ to another vertex
$V_j$ is $\alpha/d_{out}(V_i) + (1-\alpha)/|V|$, where $\alpha$ is a
user-defined parameter (usually 0.8-0.9), $d_{out}(V_i)$ is the
outdegree of vertex $V_i$, and $|V|$ is the cardinality of $V$.  The
first term represents the probability of following a given link on
page $V_i$; the second represents the probability of moving to a
random page.  For pages with no outlinks, the first term is
$\alpha/|V|$, indicating equal likelihood to move to any other page.
Equivalently, the graph can be represented by a matrix
$A$~\cite{LangvilleMeyer05a}, with matrix entries $A_{ij} =
\alpha/d_{out}(V_i)$ if vertex $V_i$ links to $V_j$.  To maintain the
sparsity of $A$, terms for random jumps and zero out-degree vertices
are not explicitly stored as part of the matrix.  Rather, they are
computed separately as adjustments to the PageRank vector.  The kernel
of the PageRank algorithm is thus a power-method iteration consisting
of matrix-vector multiplications $A^T x=y$, where $x$ is the PageRank
vector from the previous iteration.  A few additional dot product and
norm computations per iteration are also required.

The algorithm in Figure~\ref{fig:pr2} performs these iterations, using
linear-algebra concepts implemented in a MapReduce framework with data
stored in 4 MapReduce objects: $A$, $MT$, $x$, and $y$.  $A$ stores
the matrix, $MT$ its ``empty'' rows, while $x$ and $y$ are vectors.
The nonzeros of matrix $A$ are initialized as described above; the
PageRank vector $x$ is initialized uniformly.  A MapReduce object $MT$
contains the indices of all-zero rows of $A$, corresponding to
vertices with no outlinks.  The PageRank algorithm is made more
efficient by calling {\it aggregate()} once on the MapReduce objects
representing the matrix $A$, the all-zero matrix rows $MT$, and the
vector $x$ before the PageRank iterations begin.  Since these three
MapReduce objects have the same key types, pre-aggregating the objects
moves all KV pairs with a given key to a single processor.  Thus, many
of the PageRank operations become local operations, replacing {\it
collate()} operations (which are equivalent to an {\it aggregate()}
followed by {\it convert()}) with simply {\it convert()} calls.  In
Figure~\ref{fig:pr2}, steps where {\it convert()} operations were
substituted for {\it collate()} operations are marked with an
asterisk.  This strategy of pre-aggregating to improve data locality
is useful in many MapReduce algorithms, but assumes that specific keys
can always be mapped to the same processor.

In Step 1 of a PageRank interation, $MT$ is used in dot products to
compute adjustments to the PageRank vector to represent the uniform
probability of going from a leaf page to any other page.  Adjustments
for random jumps from any page are also computed.  In Step 2, the
matrix-vector product $A^T x$ is computed and the result storey in
$y$.  This is the most expensive part of the computation, requiring
two passes over the nonzero structure of $A$.  In the first pass, a
{\it convert()} gathers all local row entries $A_{ij}$ with their
associated $x_i$ entry, and a {\it reduce()} computes $A_{ij} x_i$.
In the second pass, a {\it collate()} gathers, for each $j$, all
contributions to the column sum $\sum_i A_{ij} x_i$, which is computed
by a second {\it reduce()}.  Steps 3 and 4 adjust and scale the
product vector; the work is proportional to $|V|$.  Step 5 computes
the residual to determine whether the PageRank vector has converged;
work is again proportional to $|V|$.  In Step 6, the PageRank vector
is overwritten prior to the next iteration.  Throughout the algorithm,
we exploit the ability to call MPI\_Allreduce to compute global norms
and residuals.

This linear-algebra approach is attractive because it allows
construction of the MapReduce PageRank algorithm from simple
operations (matrix-vector multiplication, dot products) on commonly
used data types (matrices, vectors).  One can build a variety of
algorithms based on these primitives; indeed, linear-algebra toolkits
such as Trilinos~\cite{Trilinos-Overview} and
PETSc~\cite{petsc-efficient} exploit this idea.  In parallel, these
toolkits share pre-computed data layouts and communication maps among
distributed objects to enable efficient operations between the
objects.  For MapReduce, however, separating the data into vectors and
matrices results in additional overhead to gather KV pairs with the
same key into a key-multivalue (KMV) pair.  This overhead can be
reduced in PageRank, for example, by storing the $x$ and $y$ vectors
in a single MapReduce object, with key $i$ and value $(x_i, y_i)$.
Doing so eliminates the {\it copy()} operation in Step 2 of
Figure~\ref{fig:pr2}, as well as the {\it add()} and {\it convert()}
operations in Step 5.  A further performance improvement would store
an entire row of $A$ along with $x$ and $y$ in a single MapReduce
object, with key $i$ and value $(x_i, y_i, d_{out}, [j_1, A_{ij_1}],
[j_2, A_{ij_2}], \dots, [j_{d_{out}}, A_ij_{d_{out}}])$ (that is,
storing together, for each vertex, its old and new PageRank value
along with all out-going edges incident to the vertex).  This data
layout eliminates an additional {it convert()} operation, but at the
cost of code reuse and scalability.  While this data layout may be
efficient for PageRank, it is more complex and less likely to be
useful for additional algorithms.  In addition, since each KV pair
must fit within a single page in MR-MPI, storing entire matrix rows in
a single KV pair limits the size of the the matrices on which the
algorithm can be applied.  The more general approach of storing
non-zeros separately enables arbitrarily large problems to be solved.
The detailed timing break-downs in
section~\ref{subsec:results_pagerank} allow one to estimate the
benefit of these implementations relative to code complexity and
scalability issues.

%\begin{figure}[htb]
% \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
%  xxxx\=xxxxx\=xxxxx\=xxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill

%\> MapReduce object $A$ containing $|V| \times |V|$ matrix:  Key = $i$, Value = $[j, A_{ij}]$ \\
%\> MapReduce object $MT$ containing indices of all-zero rows of $A$:  Key = $i$, Value = 0 \\
%\> MapReduce object $x$ containing initial PageRank vector:  Key = $i$, Value = $1/|V|$ \\
%While (${residual} > {tolerance}$) \\
%\> 1. Compute contribution for random jumps and zero-outdegree vertices \\
%\> \> Add: $x$ to $MT$ \\
%\> \> Collate $MT$:  Vertex is Key \\
%\> \> Reduce $MT$:   \\
%\> \> \> Input:  Key = $i$; MultiValue = $x_i, 0$ if $i \in {MT}$; MultiValue = $x_i$ if $i \notin {MT}$ \\
%\> \> \> Compute:  Local contributions $c_{local}$ \\
%\> \> \> Output:  If ${nvalues} = 2$, Key = $i$; Value = 0 \\
%\> \> MPI\_Allreduce $c_{local}$ with MPI\_SUM to compute global contribution $c_{global}$ \\
%\> 2. Compute $y = A^T x$ \\
%\> \> Copy $x$ to $y$ \\
%\> \> Add $A$ to $y$ \\
%\> \> Collate $y$:  Key is row index of $A$ \\
%\> \> Reduce $y$: \\
%\> \> \>  Input: Key = $i$; Multivalue = $i, [j, A_{ij}]$ for nonzeros $A_{ij}$ in $A$ \\
%\> \> \>  Output:  Key = $j$; Value = $A_{ij} y_i$ \\
%\> \> Collate $y$:  Key is column index of $A$ \\
%\> \> Reduce $y$:  \\
%\> \> \> Input:  Key = $j$; Multivalue = $A_{ij} y_i$ for all non-zero entries $A_{ij}$ of column $j$ \\
%\> \> \> Output:  Key = $j$; Value = $\sum_i A_{ij} y_i$ \\
%\> 3. Add contribution $c_{global}$ to $y$; compute local max norm $n_{local}$ of $y$ \\
%\> \> Map $y$:  \\
%\> \> \> Input:  Key = $i$, Value = $y_i$ \\
%\> \> \> Compute: $y_i = y_i + c_{global}$; if $y_i > n_{local}$, $n_{local} = y_i$ \\
%\> \> \> Output:  Key = $i$, Value = $y_i$ \\
%\> \> MPI\_Allreduce $n_{local}$ with MPI\_MAX to compute global max norm $n_{global}$ \\
%\> 4. Scale $y$ by $n_{global}$ \\
%\> \> Map $y$: \\
%\> \> \> Input:  Key = $i$, Value = $y_i$ \\
%\> \> \> Output:  Key = $i$, Value = $y_i / {n_{global}}$ \\
%\> 5. Compute residual \\
%\> \> Add $y$ to $x$ \\
%\> \> Collate $x$:  Key is index $i$ \\
%\> \> Reduce $x$: \\
%\> \> \> Input:  Key = $i$; Multivalue = $x_i, y_i$ \\
%\> \> \> Compute:  if $|x_i - y_i| > {resid}$, ${resid} = |x_i - y_i|$ \\
%\> \> MPI\_Allreduce with MPI\_MAX to compute $residual$ \\
%\> 6. $x = y$; \\
%Output:  PageRank vector $x$
  %\end{tabbing}
 %\end{minipage}}\end{center}
%
 %\caption{\it MapReduce algorithm for PageRank vertex ranking, using
 %4 MapReduce objects $A$, $MT$, $x$, and $y$.}
%
 %\label{fig:pr}
%\end{figure}

\begin{figure}[htb]
 \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
  xxxx\=xxxxx\=xxxxx\=xxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill

$*$ Pre-aggregated MapReduce object $A$ = $|V| \times |V|$ matrix:  Key = $i$, Value = $[j, A_{ij}]$ \\
$*$ Pre-aggregated MapReduce object $MT$ = indices of all-zero rows of $A$:  Key = $i$, Value = 0 \\
$*$ Pre-aggregated MapReduce object $x$ = initial PageRank vector:  Key = $i$, Value = $1/|V|$ \\

While (${residual} > {tolerance}$) \\
\> 1 Compute contribution for random jumps and zero-outdegree vertices \\
\> \> Add: $x$ to $MT$ \\
$*$\> \> Convert $MT$:  Vertex is Key \\
\> \> Reduce $MT$:   \\
\> \> \> Input:  Key = $i$; MultiValue = $x_i, 0$ if $i \in {MT}$; MultiValue = $x_i$ if $i \notin {MT}$ \\
\> \> \> Compute:  Local contributions $c_{local}$ \\
\> \> \> Output:  If ${nvalues} = 2$, Key = $i$; Value = 0 \\
\> \> MPI\_Allreduce $c_{local}$ with MPI\_SUM to compute global contribution $c_{global}$ \\
\> 2 Compute $y = A^T x$  \\
\> \> Copy $x$ to $y$ \\
\> \> Add $A$ to $y$ \\
$*$\> \> Convert $y$:  Key is row index of $A$ \\
\> \> Reduce $y$: \\
\> \> \>  Input: Key = $i$; Multivalue = $y_i, [j, A_{ij}]$ for nonzeros $A_{ij}$ in $A$ \\
\> \> \>  Output:  Key = $j$; Value = $A_{ij} y_i$ \\
\> \> Collate $y$:  Key is column index of $A$ \\
\> \> Reduce $y$:  \\
\> \> \> Input:  Key = $j$; Multivalue = $A_{ij} y_i$ for all non-zero entries $A_{ij}$ of column $j$ \\
\> \> \> Output:  Key = $j$; Value = $\sum_i A_{ij} y_i$ \\
\> 3 Add contribution $c_{global}$ to $y$; compute local max norm $n_{local}$ of $y$ \\
\> \> Map $y$:  \\
\> \> \> Input:  Key = $i$, Value = $y_i$ \\
\> \> \> Compute: $y_i = y_i + c_{global}$; if $y_i > {n_{local}}$, $n_{local} = y_i$ \\
\> \> \> Output:  Key = $i$, Value = $y_i$ \\
\> \> MPI\_Allreduce $n_{local}$ with MPI\_MAX to compute global max norm $n_{global}$ \\
\> 4 Scale $y$ by $n_{global}$ \\
\> \> Map $y$: \\
\> \> \> Input:  Key = $i$, Value = $y_i$ \\
\> \> \> Output:  Key = $i$, Value = $y_i / {n_{global}}$ \\
\> 5 Compute residual \\
\> \> Add $y$ to $x$ \\
$*$\> \> Convert $x$:  Key is index $i$ \\
\> \> Reduce $x$:   \\
\> \> \> Input:  Key = $i$; Multivalue = $x_i, y_i$ \\
\> \> \> Compute:  if $|x_i - y_i| > {resid}$, ${resid} = |x_i - y_i|$ \\
\> \> MPI\_Allreduce with MPI\_MAX to compute $residual$ \\
\> 6 $x = y$; \\
Output: PageRank vector $x$
  \end{tabbing}
 \end{minipage}}\end{center}

 \caption{\it MapReduce algorithm for PageRank vertex ranking, using 4
 MapReduce objects $A$, $MT$, $x$, and $y$.  Pre-aggregating several
 of the MapReduce objects allows many communication intensive {\it
 collate()} operations to be replaced by local {\it convert()}
 operations (marked with asterisks).}

 \label{fig:pr2}
\end{figure}

\subsection{Triangle enumeration}

A triangle in a graph is any triplet of vertices $(V_i,V_j,V_k)$ where
the edges $(V_i,V_j), (V_j,V_k), (V_i,V_k)$ exist.  Figure
\ref{fig:tri} outlines a MapReduce algorithm that enumerates all
triangles, assuming an input graph $G$ of undirected edges $(V_i,V_j)$
where $V_i < V_j$ for every edge, i.e. an upper-triangular R-MAT
matrix.  This exposition follows the triangle-finding algorithm
presented in~\cite{Cohen09}.

\begin{figure}[htb]
 \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
  xxxxxxxxxxxx\=xxx\=xxx\=xxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill

MapReduce object $G$ = edges of graph: Key = $(V_i,V_j)$ with $V_i < V_j$, Value = NULL \\

1 Copy: \> $G_0$ = copy of edge KV pairs from input graph \\
1 Map $G$: \> Convert edges to vertices \\
           \> \> input Key = $(V_i,V_j)$, Value = NULL \\
           \> \> output Key = $V_i$, Value = $V_j$ \\
           \> \> output Key = $V_j$, Value = $V_i$ \\
1 Collate $G$ \\
1 Reduce $G$: \> Add first degree to one vertex in edge \\
              \> \> input Key = $V_i$, MultiValue = $(V_j, V_k, ...)$ \\
	      \> \> for each V in MultiValue: \\
              \> \> \> if $V_i < V$: output Key = $(V_i,V)$, Value = $(D_i,0)$ \\
              \> \> \> else: output Key = $(V,V_i)$, Value = $(0,D_i)$ \\
2 Collate $G$ \\
2 Reduce $G$: \> Add second degree to other vertex in edge \\
              \> \> input Key = $(V_i,V_j)$, MultiValue = $((D_i,0),(0,D_j))$ \\
              \> \> output Key = $(V_i,V_j)$, Value = $(D_i,D_j)$ with $V_i < V_j$ \\
3 Map $G$: \> Low degree vertex emits edges \\
           \> \> if $D_i < D_j$: output Key = $V_i$, Value = $V_j$ \\
           \> \> else if $D_j < D_i$: output Key = $V_j$, Value = $V_i$ \\
           \> \> else: output Key = $V_i$, Value = $V_j$ \\
3 Collate $G$ \\
3 Reduce $G$: \> Emit angles of each vertex \\
              \> \> input Key = $V_i$, MultiValue = $(V_j, V_k, ...)$ \\
	      \> \> for each $V_1$ in MultiValue: \\
	      \> \> \> for each $V_2$ beyond $V_1$ in MultiValue: \\
	      \> \> \> \> if $V_1 < V_2$: output Key = $(V_1,V_2)$, Value = $V_i$ \\
	      \> \> \> \> else: output Key = $(V_2,V_1)$, Value = $V_i$ \\
4 Add: \> Add $G_0$ edge KV pairs to angle KV pairs in $G$ \\
4 Collate $G$ \\
4 Reduce $G$: \> Emit triangles \\
              \> \> input Key = $(V_i,V_j)$, MultiValue = $(V_k,V_l,NULL,V_m,...)$ \\
              \> \> if NULL exists in MultiValue: \\
	      \> \> \> for each non-NULL V in MultiValue: \\
	      \> \> \> \> output Key = $(V_i,V_j,V)$, Value = NULL

  \end{tabbing}
 \end{minipage}}\end{center}

 \caption{\it MapReduce algorithm for triangle enumeration, using 2 MapReduce objects $G$ and $G_0$}

 \label{fig:tri}
\end{figure}

The initial step is to store a copy of the graph edges as key/value
(KV) pairs in an auxiliary MapReduce object $G_0$, for use later in
the algorithm.  The first {\it map()} operation on $G$ converts edge
keys to vertex keys with edge values.  After a {\it collate()}, each
vertex has a list of vertices it is connected to; the first {\it
reduce()} can thus flag one vertex $V_i$ in each edge with a degree
count $D_i$.  The second {\it collate()} and {\it reduce()} assign a
degree count $D_j$ to the other vertex in each edge.  In the third
{\it map()}, only the lower-degree vertex in each edge emits its edges
KV pairs.  The task of the third {\it reduce()} is to emit ``angles''
for each of these low-degree vertices.  An ``angle'' is a root vertex
$V_i$, with two edges to vertices $V_1$ and $V_2$, i.e. a triangle
without the third edge $(V_1,V_2)$.  The {\it reduce()} emits a list
of all angles of vertex $V_i$, by a double loop over the edges of
$V_i$.  Note that the aggregate volume of KV pairs emitted at this
stage is minimized by having only the low-degree vertex in each edge
generate angles.

In stage 4, KV pairs from the original graph $G_0$ are added to the
current working set of KV pairs in MapReduce object $G$.  The KV pairs
in $G_0$ are edges that complete triangles for the angle KV pairs just
generated.  After the fourth {\it collate()}, a pair of vertices
$(V_i,V_j)$ is the key, and the multivalue is the list of all root
vertices in angles that contain $V_i$ and $V_j$.  If the multivalue
also contains a NULL, contributed by $G_0$, then there is a
$(V_i,V_j)$ edge in the graph.  Thus all vertices in the multivalue
are roots of angles that are complete triangles and can be emitted as
a triplet key.

\subsection{Connected component labeling}

A connected component of a graph is a set of vertices where all pairs
of vertices in the set are connected by a path of edges.  A sparse
graph may contain many such components.  Figure~\ref{fig:cc} outlines
a MapReduce algorithm that labels each vertex in a graph with a
component ID.  All vertices in the same component are labelled with
the same ID, which is the ID of some vertex in the component.  We
assume an input graph $E$ of undirected edges $(V_i,V_j)$.  This
exposition also follows the connected-component algorithm presented
in~\cite{Cohen09}, with the addition of logic that load-balances data
across processors, which is important when one or a few giant
components exist in the graph.

\begin{figure}[htb]
 \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
  xxxx\=xxxxxxxxxxx\=xxxx\=xxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill

MapReduce object $E$ = edges of graph: Key = $(V_i,V_j)$, Value = NULL \\
MapReduce object $V$ = zone assignment of each vertex: Key = $V_i$, Value = $Z_i$ \\
MapReduce object $W$ = workspace

Iterate: \\

\> 1 Map $W$ using $E$ as input: \> Convert edges to vertices \\
    \> \> \> input Key = $E_{ij} = (V_i, V_j)$, Value = NULL \\
    \> \> \> output Key = $V_i$, Value = $E_{ij}$ \\
    \> \> \> output Key = $V_j$, Value = $E_{ij}$ \\
\> 1 Add $V$ to $W$ \\
\> 1 Collate $W$: \> Vertex as key \\
\> 1 Reduce $W$: \> Emit edges of each vertex with zone of vertex \\
       \> \> \> input Key = $V_i$, MultiValue = $E E E E ... Z$ \\
       \> \> \> for each $E$ in MultiValue: \\
      \> \> \> \> output Key = $E_{ij}$, Value = $Z_i$ \\

\> 2 Collate $W$: \> Edge as key \\
\> 2 Reduce $W$: \> Emit zone re-assignments \\
       \> \> \> input Key = $E_{ij}$, MultiValue = $Z_i Z_j$ \\
       \> \> \> $Z_{winner}$ = min($Z_i$,$Z_j$); $Z_{loser}$ = max($Z_i$,$Z_j$) \\
       \> \> \> if $Z_i$ and $Z_j$ are different: \\
      \> \> \> \> output Key = $Z_{loser}$, Value = $Z_{winner}$ \\
\> 2 Exit: \> if $W$ is empty, i.e. no output by Reduce 2 \\
\> 3 Map $V$: \> Invert vertex/zone pairs \\
    \> \> \> input Key = $V_i$, Value = $Z_i$ \\
    \> \> \> if $Z_i$ is not partitioned: \\
   \> \> \> \> output Key = $Z_i$, Value = $V_i$ \\
    \> \> \> else: \\
   \> \> \> \> output Key = $Z_i^+$, Value = $V_i$ for a random processor \\
\> 4 Map $V$ using $W$ as input: \> Add zone reassignments in $W$ to $V$ \\
    \> \> \> if $Z_i$ is not partitioned: \\
   \> \> \> \> output Key = $Z_i$, Value = $Z_{winner}$ \\
    \> \> \> else: \\
   \> \> \> \> output Key = $Z_i^+$, Value = $Z_{winner}$ for every processor \\
   \> \> \> \> output Key = $Z_i$, Value = $Z_{winner}$ \\
\> 4 Collate $V$: \> Zone ID as key \\
\> 4 Reduce $V$: \> Emit new zone assignment of each vertex \\
       \> \> \> input Key = $Z_i$ or $Z_i^+$, MultiValue = $V V V V ... Z Z Z ...$ \\
       \> \> \> $Z_{new}$ = min($Z_i$ or $Z_i^+$,Z,Z,Z,...) \\
       \> \> \> partition $Z_{new}$ if number of $V >$ threshhold \\
       \> \> \> for each $V$ in MultiValue: \\
      \> \> \> \> output Key = $V_i$, Value = $Z_{new}$

  \end{tabbing}
 \end{minipage}}\end{center}

 \caption{\it MapReduce algorithm for connected component labeling,
 using 3 MpaReduce objects $E$, $V$, and $W$.}

 \label{fig:cc}
\end{figure}

The algorithm begins (before the iteration loop) by assigning each
vertex to its own component or ``zone,'' so that $Z_i$ = $V_i$ and
storing this assignment in MapReduce object $V$.  Each iteration grows
the zones, one layer of neighbors at a time.  As zones collide due to
shared edges, a winner is chosen (the smaller zone ID), and vertices
in the losing zone are reassigned to the winning zone.  When the
iterations complete, each zone has become a fully connected component
and the MR object $V$ holds that zone assignment for every vertex.
The algorithm thus finds all connected components in the graph
simultaneously.  The number of iterations required depends on the
largest diameter of any component in the graph.

The first {\it map()} operation uses MapReduce object $E$ as input to
the empty MapReduce object $W$ to emit the vertices in each edge as
keys, with the edge as a value.  The current zone assignment of each
vertex in $V$ is added to the set of key/value (KV) pairs in $W$.  The
first {\it collate()} operation collects all the edges of a vertex and
its zone assignment together in one multi-value.  The first {\it
reduce()} operation re-emits each edge, tagged by the zone assignment
of one of its vertices.

Since each edge was emitted twice, the second {\it collate()}
operation on $W$ collects the zone assignments for its two vertices
together.  If the two zone IDs are different, the second {\it
reduce()} operation chooses a winner (the smaller of the two IDs), and
emits the loser ID as a key, with the winning ID as a value.  If no
zone ID changes are emitted, the algorithm is finished, and the
iterations cease.

The third {\it map()} operation inverts the vertex/zone KV pairs to
become zone/vertex KV pairs.  The partitioning logic described in
Figure~\ref{fig:cc} for this step (and subsequent ones) is described
below.  The fourth {\it map()} operation adds the changed zone
assignments in $W$ to the set of KV pairs in $V$.  The fourth {\it
collate()} can now collect all the vertices in each zone along with
zero or more reassignments for that zone ID.  Since a zone could
collide with multiple other zones on the same iteration due to shared
edges, the new zone ID for a zone becomes the minimum ID of any of its
neighboring zones.  If no zone reassignment values appear in the
multi-value resulting from the fourth {\it collate()}, the zone ID is
unchanged.  The final {\it reduce()} emits a KV pair for each vertex
in the zone, with the vertex as a key and the new zone ID as a value.
This is updated version of MapReduce object $V$, ready to be used as
input for the next iteration.

Note that if a graph has only a few components, the fourth {\it
collate()} operation, which keys on the zone ID, may generate a few
very large key/multi-value (KMV) pairs.  For example, if the entire
graph is fully connected, in the last iteration, a single KMV pair
assigned to one processor will result, containing all vertices in the
graph.  This imbalance in memory and computational work can lead to
poor parallel performance of the overall algorithm.  To counter this
effect, the various operations in stages 3 and 4 include extra logic.
The idea is to partition zones whose vertex count exceeds a
user-defined threshhold into $P$ sub-zones, where $P$ is the number of
processors.  The 64-bit integer that stores the zone ID also stores a
bit flag indicating the zone has been partitioned and a set of bits
that encode the processor ID.

During the third {\it map()} operation, if the zone has been
partitioned, the vertex is assigned to a random processor and the
processor ID bits are added to the zone ID, as indicated by the
$Z_i^+$ notation in Figure~\ref{fig:cc}.  Likewise, if $Z_i$ has been
partitioned, the fourth {\it map()} operation emits the KV pair for
the zone ID as ($Z_i^+$,$Z_{winner}$) instead of as
($Z_i$,$Z_{winner}$).  In this case ($Z_i$,$Z_{winner}$), is emitted
not once, but $P+1$ times, once for each processor, and once as if
$Z_i$ had not been partitioned (for a reason discussed below).

With this additional partitioning logic, the fourth {\it collate()}
operation (which keys on zone IDs, some of which now include processor
bits) collects only $1/P$ of the vertices in large zones onto each
processor.  Yet the multivalue on each processor still contains all
the zone-reassignments relevant to the unpartitioned zone.  Thus, the
fourth {\it reduce()} operation can change the zone ID (if necessary)
in a consistent manner across all $P$ multi-values that contain the
zone's vertices.  When the fourth {\it reduce()} operation emits new
zone assignments for each vertex, the zone retains its partitioned
status; the partition bit is also explicitly set if the vertex count
exceeds the threshhold for the first time.

Note that this logic does not guarantee that the partition bits of the
zone IDs for all the vertices in a single zone are set consistently on
a given iteration.  For example, an unpartitioned zone with a small ID
may consume a partitioned zone.  The vertices from the partitioned
zone retain their partitioned status, but the original vertices in the
small zone may not set the partition bit of their zone IDs.  On
subsequent iterations, the fourth {\it map()} operation emits $P+1$
copies of new zone reassignments for both partitioned and unpartioned
zone IDs, to ensure all vertices in the zone know the correct
reassignment information.

\subsection{Maximal independent set identification}

An ``independent'' set of vertices from a graph is one where no pair
of vertices in the set is connected by an edge.  The set is
``maximal'' if no vertex can be added to it \footnote{A maximal set is
``maximum'' if it is the largest maximal set.  Finding a graph's
maximum independent set is an NP-hard problem, which MapReduce is
unlikely to help with.}.  Finding a maximal independent set (MIS) is
useful in several contexts, such as identifying large numbers of
independent starting vertices for graph traversals.  Luby's
algorithm~\cite{Luby86} is a well-known parallel method for finding a
MIS.  Figure~\ref{fig:luby} outlines a MapReduce version of Luby's
algorithm, assuming an input graph of undirected edges $(V_i,V_j)$
where $V_i < V_j$ for every edge, i.e. an upper-triangular R-MAT
matrix.

\begin{figure}[htb]
 \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
  xxxx\=xxxxxxxxxxx\=xxxx\=xxxx\=xxxx\=xxxxxxxxxxxxxxxxxxxxxxxxx\= \kill

MapReduce object $V$ = maximal independent set vertices, initially empty \\
MapReduce object $E$ = edges of graph: Key = $(V_i,V_j)$ with $V_i < V_j$, Value = NULL \\
Map $E$: assign random values to each vertex $V_i$ and $V_j$ in $E$ \\
Clone $E$: convert KV pairs in $E$ directly to KMV pairs with multivalue = NULL \\
\\
while $N_{edges}$ in $E$ > 0: \\
\> 1 Reduce $E$: \> Determine WINNER/LOSER vertex of each edge \\
       \> \> \> input Key = $E_{ij}$, Multivalue = NULL or LOSER \\
       \> \> \> if multivalue = LOSER, emit nothing \\
       \> \> \> $V_w$ = WINNER vertex, $V_l$ = LOSER vertex \\
       \> \> \> output Key = $V_w$, Value = $V_l$, WINNER \\
       \> \> \> output Key = $V_l$, Value = $V_w$, LOSER \\
\> 1 Collate $E$: \> Vertex as key \\
\> 2 Reduce $E$: \> Find WINNER vertices \\
       \> \> \> input Key = $V_i$, MultiValue = $V V V V ...$ \\
       \> \> \> if all $V$ are flagged with WINNER, $V_i$ is a WINNER \\
      \> \> \> \> for each $V_j$ in Multivalue: \\
     \> \> \> \> \> output Key = $V_j$, Value = $V_i$, WINNER \\
       \> \> \> else: \\
      \> \> \> \> for each $V_j$ in Multivalue: \\
     \> \> \> \> \> output Key = $V_j$, Value = $V_i$ \\
\> 2 Collate $E$: \> Vertex as key \\
\> 3 Reduce $E$: \> Find LOSER vertices \\
       \> \> \> input Key = $V_i$, MultiValue = $V V V V ...$ \\
       \> \> \> if any $V$ is flagged with WINNER, $V_i$ is a LOSER \\
      \> \> \> \> for each $V_j$ in Multivalue: \\
     \> \> \> \> \> output Key = $V_j$, Value = $V_i$, LOSER \\
       \> \> \> else: \\
      \> \> \> \> for each $V_j$ in Multivalue: \\
     \> \> \> \> \> output Key = $V_j$, Value = $V_i$ \\
\> 3 Collate $E$: \> Vertex as key \\
\> 4 Reduce $E$: \> Emit WINNER vertices and LOSER edges \\
       \> \> \> input Key = $V_i$, MultiValue = $V V V V ...$ \\
       \> \> \> if all $V$ are flagged with LOSER, $V_i$ is a WINNER \\
      \> \> \> \> output to $V$: Key = $V_i$, Value = NULL \\
       \> \> \> for each $V_j$ in Multivalue: \\
      \> \> \> \> if $V_j$ is flagged with LOSER: \\
     \> \> \> \> \> output to $E$: Key = $E_{ij}$ with $V_i < V_j$, Value = LOSER \\
      \> \> \> \> else: \\
     \> \> \> \> \> output to $E$: Key = $E_{ij}$ with $V_i < V_j$, Value = NULL \\
\> 4 Collate $E$: \> Edge as key \\

  \end{tabbing}
 \end{minipage}}\end{center}

 \caption{\it MapReduce algorithm for finding a maximal independent
 set of graph vertices via Luby's algorithm, using two MapReduce
 objects $E$ and $V$.}

 \label{fig:luby}
\end{figure}

Before the iterations begin, a random value is assigned to each vertex
(stored consistently in multiple edges in $E$), which is used to
compare pairs of vertices.  The vertex ID itself could be used as the
random value, but since the vertices eventually selected for the MIS
depend on the randomization, this may introduce an unwanted bias.
Instead a random number generator can be used to assign a consistent
random value $R_i$ to each vertex in each edge (via a {\it map()}
operation), which is then carried along with the vertex ID through
each stage of the algorithm.  In the notation of
Figure~\ref{fig:luby}, each $V_i$ is then really two quantities, a
vertex ID (1 to $N$) and $R_i$.  Alternatively, the $R_i$ can be used
to relabel the vertices in a random manner, assigning each a new,
unique vertex ID from 1 to $N$.  In this case, $V_i$ is simply the new
vertex ID, and the vertices in the final MIS could be remapped to
recover their original IDs.  These operations can be performed with a
few extra MapReduce steps, outside the iteration loop.

The pre-iteration {\it clone()} operation converts the initial list of
edge key/value (KV) pairs stored in MapReduce object $E$ to
key/multivalue (KMV) pairs, with the edge as the key, and NULL as the
value.  Thus, the iterations can begin with a {\it reduce()}
operation, without need for a {\it collate()} operation.  A second
empty MapReduce object $V$ is also created, which will accumulate MIS
vertices as the iterations proceed.

At each iteration, the current set of edges in $E$ is examined to
identify winning vertices.  Vertices adjacent to winners are flagged
as losers, and all edges which contain a loser vertex are marked for
removal at the start of the next iteration.  Vertices with all their
edges marked as losers are also flagged as winners.  At each iteration
marked edges are removed from the graph, and winning vertices are
added to the MIS.  When the graph is empty, the MIS is complete.

The first {\it reduce()} operation flags the two vertices in each edge
as a potential winner or loser.  This is done by comparing the two
random values for the vertices (or the randomized vertex IDs as
explained above).  The first {\it collate()} operation thus produces a
multivalue for each vertex $V_i$ which contains all the vertices it
shares an edge with, and associated winner/loser flags.

In the second {\it reduce()} operation, if $V_i$ won the comparison
with all its neighbor vertices, then it becomes a ``winner'' vertex.
All vertices connected to the winner are emitted with $V_i$ tagged
with a winner flag.  If $V_i$ is not an overall winner, its neighbor
vertices are emitted without the winner flag on $V_i$..  The second
{\it collate()} operation collects this new information for each
vertex.

In the third {\it reduce()} operation, if any neighbor of $V_i$ was a
winner in the previoue {\it reduce()} operation (indicated by a winner
flag), $V_i$ becomes a losing vertex and emits all its neighbor
vertices with $V_i$ tagged with a loser flag.  Otherwise it emits all
its neighbor vertices without the loser flag on $V_i$.

In the fourth {\it reduce()} operation, some vertices have all their
neighbor vertices flagged as losers; these vertices become winners.
They are either the vertices identified as winners in the second {\it
reduce()} operation, or vertices who would become singletons (having
no edges) when edges containing a loser vertex are removed (on the
next iteration).  These winning vertices are emitted to the MapReduce
object $V$ that stores the accumulating MIS.

All edges of each vertex are also emitted, retaining the loser flag if
they have it; they are emitted with $E_{ij} = (V_i, V_j)$ as their
key, with $V_i < V_j$.  This is to ensure both copies of the edge come
together via the fourth {\it collate()} operation.  The multivalue for
each edge now has two values, each of which is either NULL or a loser
flag.

In the first {\it reduce()} of the next iteration, any edge with a
loser flag in its multivalue does not compare its two vertices; it is
effectively deleted from the graph.  The Luby iterations end when all
edges have been removed from the graph.  At this point, the MapReduce
object $V$ contains a complete MIS of vertices.

\subsection{Single source shortest path calculation}

For a directed graph with weighted edges, the single-source shortest
path (SSSP) algorithm computes the shortest weighted distance from a
chosen source vertex to all other vertices in the graph.  This
calculation is straightforward when global graph information is
available.  The source vertex is labeled with distance zero.  Then in
each iteration, edge adjacencies of labeled vertices are followed, and
adjacent vertices are relabeled with updated distances.  When no
distances change, the iterations are complete.  Most importantly, only
edges adjacent to modified vertices need be visited in each iteration.

For a distributed graph, global information about edge adjacencies and
labeled vertices is not available.  Instead, following the
Bellman-Ford-style algorithm of~\cite{SSSPMapReduce, Bellman58,
Ford62}, every edge of the graph is visited in each iteration and
vertex distancs are updated.  A MapReduce formulation of this
algorithm, described in Figure~\ref{fig:sssp}, stores directed edges
in $E$ and begins by labeling the distance to all vertices as
inifinte, except a distance of 0.0 to the source vertex $V_s$.  At
each iteration, a {\it map()} operation is performed using $E$ as
input to add edges and their weights to each vertex in $V$.  After a
{\it collate()} operation, each vertex $V_i$ knows its outgoing edges
and their weights, as well as its own current distance.  It may also
have one or more distances assigned to it from a previous iteration
(i.e. from an inbound edge).  In the {\it reduce()} operation, each
vertex resets its distance to the minimum of these new distances and
its current distance.  If its distance changed, it emits new distance
values $d_j$ for the vertices associated with its outgoing edges,
where $d_j = d_i + w_{ij}$.  It also unsets a ``done'' flag, stored by
each processor.  When the reduce is complete, an MPI\_Allreduce
operation is performed to check the ``done'' flag status on all
processors.  If any distance was changed, the iterations continue, and
the distance updates propagate through the graph.  If no distances
changed, the algorithm is done.

\begin{figure}[htb]
 \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
  xxxx\=xxxx\=xxxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill

MapReduce object $V$ = vertices with distances from source vertex $V_s$: \\
\> Key = $V_i$, Value = $\infty$, except Value for $V_s$ = 0 \\
MapReduce object $E$ = graph edges with weights: Key = ($V_i, V_j$), Value = $w_{ij}$ \\
$done$ = 0 \\

while not $done$: \\
\> $done$ = 1 \\
\> Map $V$ using $E$ as input: add edge weights in $E$ to $V$ \\
\> \> input Key = $E_{ij}$, Value = $w_{ij}$ \\
\> \> output Key = $V_i$, Value = $[V_j, w_{ij}]$ \\
\> Collate $V$: Vertex is key \\
\> Reduce $V$:  Loop over edges of vertex to update distance to $V_s$ \\
\> \> Input: Key = $V_i$; Multivalue = edges $[V_j, w_{ij}]$ and candidate distances for $d_i$ \\
\> \> Find smallest new distance $d_{min}$ for $V_i$ \\
\> \> If $d_{min} \ne d_i$: \\
\> \> \>  $done$ = 0 \\
\> \> \>  Output: Key = $V_j$, Value = $d_{min} + w_{ij}$ for each edge vertex $V_j$ \\
\> \>  Output:  Key = $V_i$, Value = $d_{min}$ \\
\> $done$ = MPI\_Allreduce($done$,MPI\_MIN)

  \end{tabbing}
 \end{minipage}}\end{center}

 \caption{\it MapReduce algorithm for single-source shortest path
 (SSSP) calculation, using two MapReduce objects $V$ and $E$.}

 \label{fig:sssp}
\end{figure}

As with PageRank and R-MAT generation, significant savings in
communication and execution time can be achieved by pre-aggregating
the data stored in MapReduce objects and by splitting the data
appropriately across more than one MapReduce object.  An enhanced SSSP
algorithm is shown in Figure~\ref{fig:sssp2}.  The MapReduce objects
$V$ and $E$ which store vertices and edges are aggregated across
processors only once at the beginning of the computation, which means
the communication portion of the {\it collate()} operation is
performed to assign each key to a processor.  The {\it collate()}
operation of Figure~\ref{fig:sssp} can then be replaced with a {\it
compress()} operation which requires no communication.

In each iteration of SSSP, typically only a small number of new
distances are generated.  The enhanced algorithm uses MapReduce object
$U$ to store these updates.  Only the updates in $U$ are communicated;
which aggregates them to the same processor that permanently stores
the corresponding vertex and its edges.  Thus, the total amount of
required communication is smaller than in the original algorithm,
which communicated all graph edges at each iteration.  Updated vertex
distances are stored in MapReduce object $V$, as before.

\begin{figure}[htb]
 \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
  xxxx\=xxxx\=xxxx\=xxxx\=xxxx\=xxxx\=xxxxxxxxxxxxxxxxxxxxxx\= \kill

Pre-aggregated MapReduce object $V$ = graph vertices $V_i$ with distances from sourece vertex $V_s$:  \\
\> Key = $V_i$; Value = $\infty$, except Value for $V_s$ = 0 \\
Pre-aggregated MapReduce object $E$ = graph edges with weights: Key = ($V_i, V_j$), Value = $w_{ij}$ \\
MapReduce object $U$ to store updates to distances, initially empty \\
done = 0 \\

while not $done$: \\
\> $done$ = 1 \\
\> Aggregate $U$ \\
\> Add $U$ to $V$, empty $U$ \\
\> Compress $V$:  Compute best distance for each vertex \\
\> \> Input:  Key = $V_i$; Multivalue = candidate distances for $d_i$ \\
\> \> Find smallest new distance $d_{min}$ for $V_i$ \\
\> \> If $d_{min} \ne d_i$: \\
\> \> \> $done$ = 0 \\
\> \> \> Output to $U$: Key = $V_i$; Value = $d_{min}$ \\
\> \> Output to $V$: Key = $V_i$; Value = $d_{min}$ \\
\> $done$ = MPI\_Allreduce($done$,MPI\_MIN) \\
\> if not $done$: \\
\> \> Add $U$ to $E$, empty $U$ \\
\> \> Compress $E$: Generate candidate distances for neighbors of changed vertices. \\
\> \> \> Input:  Key = $V_i$; Multivalue = $[V_j, w_{ij}]$ and candidate distance $d_i$ if updated in previous step \\
\> \> \> If updated distance $d_i$ exists: \\
\> \> \> \> Output to $U$: Key = $V_j$, Value = $d_i+w_{ij}$ for each edge vertex $V_j$) \\

  \end{tabbing}
 \end{minipage}}\end{center}

 \caption{\it Enhanced MapReduce algorithm for single-source shortest
 path (SSSP) calculation, with an additional MapReduce object $U$.
 Pre-aggregating the vertices $V$ and edges $E$, and storing only
 changed distances in $U$, reduces communication and execution time.}

 \label{fig:sssp2}
\end{figure}