### 2010-11-12 paper/results.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@665 3fc257bb-8550-0410-81b3-a5c103da83f7

\section{Performance Results}
\label{sec:results}

In this section, we present performance results for the graph
algorithms of the preceeding section, implemented as small C++
programs calling our MR-MPI library.  The benchmarks were run on a
medium-sized Linux cluster of 2 GHz dual-core AMD Opteron processors
connected via a Myrinet network.  Most importantly for MR-MPI, each
node of the cluster has one local disk, which is used for out-of-core
operations.  To avoid contention for disk I/O, we ran all experiments
with one MPI process per node.  For comparisons with other
implementations, we used either the same cluster or, where noted,
Sandia's Cray XMT, a multi-threaded parallel computer with 500 {MHz}
processors and a 3D-Torus network.

We ran the algorithms on three R-MAT graphs of different sizes, each
for a varying number of processors.  Details of the input data are
shown in Table~\ref{t:rmats}.  The {\it small} problem (around 8M
edges) can typically be run on a single processor without incurring
out-of-core operations.  The {\it medium} problem (around 134M edges)
can be run on a single processor with out-of-core operations; larger
processor configurations, however, do not necessarily require
out-of-core operations.  The {\it large} problem (around 2B edges)
requires out-of-core operations on our 64-node cluster.  These are
edge counts before pre-processing steps used in some of the
algorithms, e.g. to eliminate self-edges or convert the matrix to
upper-triangular form.

%The {\it x-large} problem 
%(around 34B edges) requires most of the machine to run.

All data sets used R-MAT parameters $(a, b, c, d) = (0.57, 0.19, 0.19,
0.05)$ and generated 8 edges per vertex (on average).  These
parameters create a highly skewed degree distribution, as indicated by
the maximumn vertex degree in Table~\ref{t:rmats}.

\begin{table}
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|c|c|}
\hline
Data & \# of    & \# of & Maximum \\
Set  & vertices & edges & vertex degree\\
\hline
RMAT-20 (small)   &$2^{20} \approx 1M$ & $2^{23} \approx 8M$ &  $\approx 24K$ \\
RMAT-24 (medium)  &$2^{24} \approx 17M$ & $2^{27} \approx 134M$ &  $\approx 147K$ \\
RMAT-28 (large)   &$2^{28} \approx 268M$ & $2^{31} \approx 2B$& $\approx 880K$ \\
%RMAT-32 (x-large) &$2^{32} \approx 4B$ & $2^{35} \approx 34B$ &   \\
\hline
\end{tabular}
\caption{Characteristics of R-MAT input data for graph algorithm
benchmarks.}
\label{t:rmats}
\end{center}
\end{table}

The resulting timings give a sense of the inherent scalability of the
MapReduce algorithms as graph size grows on a fixed number of
processors, and of the parallel scalability for computing on a graph
of fixed size on a growing number of processors.  Where available, we
compare the MapReduce algorithm with other parallel implementations,
including more traditional distributed-memory algorithms and also
multi-threaded algorithms in the Multi-Threaded Graph Library
(MTGL)~\cite{MTGL} on the Cray XMT.  We compute parallel efficiency on
$P$ processors as $({time}_{M} \times M) / ({time}_P \times P)$ where
$M$ is the smallest number of processors on which the same graph was
run, and ${time}_I$ is the execution time required on $I$ processors.

\subsection{R-MAT generation results}

In Figure~\ref{f:rmat}, we show the scalability of the R-MAT
generation algorithm (Figure~\ref{fig:rmat2}) for RMAT-20, RMAT-24 and
RMAT-28.  For RMAT-20 and RMAT-24, the algorithm exhibits superlinear
speed-up.  This is due to the decreased amount of file I/O needed with
greater numbers of processors; with a larger total memory, a
fixed-size problem requires less I/O as processors are added.  For
RMAT-28, which requires significant out-of-core operations, parallel
efficiency ranges from 52\% to 97\%.  For these problems the enhanced
algorithm of Figure~\ref{fig:rmat2} ran about 10\% faster than the
algorithm of Figure~\ref{fig:rmat}.

%Also in Figure~\ref{f:rmat}, we compare our MR-MPI implementation with an 
%R-MAT generator in MTGL on the Cray XMT.  In the MTGL implementation, 
%threads generate edges independently.  A global hash table is used to remove
%duplicate edges.  Our MR-MPI implementation runs more quickly than the MTGL
%implementation.  The differences in execution time are greatest on 64
%processors, where MR-MPI needs the least file I/O for out-of-core operations.

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_rmat.pdf}
\caption{Performance of the MR-MPI R-MAT generation algorithm (Figure~\ref{fig:rmat2}) 
%and an MTGL implementation.
}
\label{f:rmat}
\end{figure}

\subsection{PageRank results}
\label{subsec:results_pagerank}

In Figure~\ref{f:pr}, we show the performance of the PageRank
algorithm (Figure~\ref{fig:pr2}) compared to a distributed-memory
matrix-based implementation using the linear algebra toolkit
Trilinos~\cite{Trilinos-Overview}.  The Trilinos implementation of
PageRank uses Epetra matrix/vector classes to represent the graph and
PageRank vectors and requires the matrix and associated vectors fit in
the aggregate memory of the processors used.  Rows of matrix $A$ and
the associated entries of the PageRank vector $x$ are uniquely
assigned to processors; a random permutation of the input matrix
effectively load balances the non-zero matrix entries across
processors.  Interprocessor communication gathers $x$ values for
matrix-vector multiplication and sums partial products into the $y$
vector.  Most communication is point-to-point communication, but some
global communication is needed for computing residuals and norms of
$x$ and $y$.

Figure~\ref{f:pr} shows the execution time per PageRank iteration for
R-MAT matrices RMAT-20, RMAT-24 and RMAT-28.  Converging the PageRank
iterations to tolerance $0.002$ requires five or six iterations.
Several R-MAT matrices of each size were generated for the
experiments; the average time over the set of matrices is reported
here.  The Trilinos implementations show near-perfect strong scaling
for RMAT-20 and RMAT-24.  The MR-MPI implementations also demonstrate
good strong scaling.  However, MR-MPI's execution time is more than an
order of magnitude slower than the Trilinos implementation.  The
benefit of MR-MPI's out-of-core implementation is seen, however, with
the RMAT-28 data set, which could be run on smaller processor sets
than the Trilinos implementation, which required 64 processors for
RMAT-28.

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_pagerank.pdf}
\caption{Comparison of PageRank implementations using 
MR-MPI (Figure~\ref{fig:pr2}) and 
Trilinos' matrix/vector classes on R-MAT data sets.}
\label{f:pr}
\end{figure}

A more detailed analysis of the performance of the MR-MPI PageRank
algorithm is shown below.  We perform the analysis using RMAT-20,
which is small enough to fit in memory and thus does not need
out-of-core operations, and RMAT-28, which requires out-of-core
operation in MR-MPI.  Table~\ref{table:prdetail} shows the amount of
time spent in each step of Figure~\ref{fig:pr2} for the MR-MPI
implementation and the equivalent Trilinos implementation.  For both
implementations, the matrix-vector multiplication (Step 2) is the most
costly step.  However, the MR-MPI algorithm is consistently slower
than the Trilinos implementation in each step.  This effect is more
pronounced for RMAT-28, where MR-MPI performs out-of-core operations.

\begin{table}[htb]
\begin{center}
\begin{tabular}{|l|r|r|r|r|}
\hline 
& \multicolumn{2}{|c|}{RMAT-20} & \multicolumn{2}{|c|}{RMAT-28} \\
& MR-MPI & Trilinos & MR-MPI & Trilinos  \\
\hline
Step 1 (Compute adjustment) & 0.0203 & 0.00030 & 17.53 & 0.0838 \\
Step 2 (Multiply $A^T x$) & 0.1377 & 0.00455 & 341.26 & 2.9239 \\
Step 3 (Apply adjustment) & 0.0015 & 0.00015 & 2.14 & 0.0494 \\
Step 4 (Scale $y$)  & 0.0013 & 0.00014 & 1.74 & 0.0359 \\
Step 5 (Compute residual)  & 0.0207 & 0.00014 & 21.20 & 0.0457 \\
\hline
\end{tabular}
\caption{Detailed execution times (in seconds) per PageRank iteration
for each stage of the algorithm
in Figure~\ref{fig:pr2} using MR-MPI and Trilinos.}
\label{table:prdetail}
\end{center}
\end{table}

To understand where MR-MPI spends its effort during the PageRank
computation, we break down the operation times further for the most
expensive Step 2 of the PageRank computation; the results are in
Table~\ref{table:prdetail2}.  In the table, we divide the {\it
collate()} operation into its two parts: an {\it aggregate()} that
involves parallel communication and a {\it convert()} that gathers
local key/value (KV) pairs with matching keys into a key-multivalue
(KMV) pair.  All interprocessor communication occurs in the {\it
aggregate()}.  While the time spent in {\it aggregate()} is
significant, it is small compared to the time spent in the two {\it
convert()} operations.  When out-of-core operations are required for
RMAT-28, the communication cost in the {\it aggregate()} is a smaller
percentage of total time for Step 2, as other data manipulation
operations such as {\it convert()} and {\it add()} require a
relatively larger percentage of the total time.  Further
experimentation is needed to determine whether reorganizing the
MapReduce data, as described in Section~\ref{subsec:graph_pagerank}.
would reduce the execution time of Step 2.  Organizing the data such
that both the matrix $A$ and vectors $x$ and $y$ are in the same
MapReduce object would eliminate the {\it copy()} and {\it add()}
functions, as well as the first {\it convert()}.  However, to update
the $y$ values at the completion of Step 2, either an additional {\it
convert()} or {\it aggregate()} with greater amounts of data would be
needed.

\begin{table}[htb]
\begin{center}
\begin{tabular}{|l|r|r|}
\hline 
PageRank Step 2 & RMAT-20 & RMAT-28 \\
\hline
Copy $x$ to $y$                       &  1.1\% &  0.4\%\\
Add $A$ to $y$                        &  3.0\% & 14.3\% \\
Convert $y$                           & 23.5\% & 43.6\% \\
Reduce $y$                            &  7.7\% &  5.8\% \\
Aggregate $y$ (first part of collate) & 24.9\% &  6.8\% \\
Convert $y$ (second part of collate)  & 38.2\% & 27.4\% \\
Reduce $y$                            &  1.6\% &  1.7\% \\
\hline
\end{tabular}
\caption{Percentage of execution time in each operation of Step 2 of the MR-MPI 
PageRank algorithm in Figure~\ref{fig:pr2}.}
\label{table:prdetail2}
\end{center}
\end{table}

Similarly, in Table~\ref{table:prdetail5}, we show the execution times
required for each operation in Step 5 of Figure~\ref{fig:pr2}, for the
residual computation.  Clearly the time to convert KV pairs to KMV
pairs dominates this computation.  Reorganization of the vectors $x$
and $y$ into a single MapReduce object (as described in
Section~\ref{subsec:graph_pagerank}) would benefit this step of the
computation, as the {\it add()}, {\it convert()}, and {\it reduce()}
could be replaced by a single {\it map()}, with execution time
comparable to the vector-scaling operation of Step 4.

\begin{table}[htb]
\begin{center}
\begin{tabular}{|l|r|r|}
\hline 
PageRank Step 5 & RMAT-20 & RMAT-28 \\
\hline
Add $y$ to $x$ &  2.2\% &  9.8\% \\
Convert $x$    & 95.1\% & 84.9\% \\
Reduce $x$     &  2.5\% &  5.3\% \\
MPI\_Allreduce &  0.2\% &  0.0\% \\
\hline
\end{tabular}
\caption{Percentage of execution time in each operation of Step 5 of
the MR-MPI PageRank algorithm in Figure~\ref{fig:pr2}.}
\label{table:prdetail5}
\end{center}
\end{table}

%This result is
%due to two factors: ({\it i}) a higher volume of communication in the
%MapReduce implementation (which comunicates one datum per graph edge
%in each iteration), compared to the Trilinos implementation which only
%communicates a volume of data proportional to the number of vertices,
%and ({\it ii}) out-of-core operations in MR-MPI were performed because
%of the page-size restriction, while all Trilinos operations were
%performed in-core.  
%

\subsection{Triangle finding results}

In Figure~\ref{f:tri}, we show the performance of the triangle finding
algorithm (Figure~\ref{fig:tri}).  Execution times for this algorithm
were too large to allow the problem to be easily run with RMAT-28 on
our cluster.  Parallel efficiencies for RMAT-20 ranged from 80\% to
140\%; for RMAT-24, they ranged from 50\% to 120\%.  The RMAT-20
matrix contained about 100 million triangles; the RMAT-24 contained
2.1 billion.

%Comparisons with an MTGL implementation on the 
%Cray XMT give mixed results.  When MR-MPI uses the least file I/O for 
%out-of-core operations (e.g., RMAT-20 on 32 and 64 processors), it 
%outperforms the MTGL implementation.  However, when significant out-of-core
%operations are needed (e.g., for RMAT-24), the MTGL implementation is faster.

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_tri.pdf}
\caption{Performance of the MR-MPI triangle-finding algorithm
(Figure~\ref{fig:tri}).}
\label{f:tri}
\end{figure}

\subsection{Connected Components}

In Figure~\ref{f:cc}, we show the performance of the connected
component identification algorithm (Figure~\ref{fig:cc}).  A
comparison is made with a hybrid ``Giant Connected Component''
implementation using both Trilinos and MR-MPI.  Power law graphs often
have one or two very large components and many very small components.
The hybrid algorithm exploits this feature by using an inexpensive
breadth-first search (BFS) from the vertex with highest degree to
identify the largest components one at a time, followed by a more
expensive algorithm to identify the small components in the remainder
of the graph.  In our hybrid implementation, we intially perform
matrix-vector multiplications in Trilinos to perform a BFS, finding a
handful of components that include 70\% or more of the vertices (for
these R-MAT matrices).  We then apply our MapReduce algorithm to the
remaining graph to identify the small components.  This hybrid
approach is quite effective, reducing the execution time to identify
all components in RMAT-20 and RMAT-24 by 80-99\%, as shown in
Figure~\ref{f:cc}.  For RMAT-28, where the graph is too large to fit
into memory, Trilinos cannot perform the initial BFS, so only the
MapReduce algorithm was used.  The total number of components
identified were A, B, C for the 3 increasing R-MAT sizes; the number
of iterations required by the algorithm to find them was X, Y, Z
respectively.

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_cc.pdf}
\caption{Performance of the MR-MPI connected components algorithm (Figure~\ref{fig:cc}) compared with a hybrid ``Giant Connected Component'' algorithm based
on Trilinos.}
\label{f:cc}
\end{figure}

\subsection{Maximally independent set results}

The execution times for the maximal independent set (MIS) algorithm
(Figure~\ref{fig:luby}) are shown in Figure~\ref{f:luby}.  Like the
R-MAT generation results, superlinear speed-up of the algorithm occurs
for RMAT-20 and RMAT-24, as more of the graph fits into processor
memory and less file I/O is needed.  For RMAT-28, the algorithm
requires significant out-of-core operations. In this case, parallel
efficiency is nearly perfect going from 8 to 64 processors.  The
number of vertices in the resulting MIS were A, B, C for the 3
increasing R-MAT sizes; the number of iterations required by the
algorithm to find them was X, Y, Z respectively.

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_luby.pdf}
\caption{Performance of the MR-MPI maximal independent set algorithm (Figure~\ref{fig:luby}).}
\label{f:luby}
\end{figure}

\subsection{Single-source shortest path results}

The execution times for the SSSP algorithms of Figures~\ref{fig:sssp}
and \ref{fig:sssp2} are shown in Figure~\ref{f:sssp}.  The results
show the benefit of using the enhanced algorithm, which gave at least
a 17\% (and often greater) reduction in execution time due to reduced
communication; only the updated distances are communicated throughout
most of the enhanced algorithm.  However, the execution times are
still large compared to multi-threaded implementations; for example,
Madduri et al.~\cite{Madduri07} report execution times of only 11
seconds on 40 processors for a multithreaded implementation on a Cray
MTA-2 (the predecessor of the XMT) on graphs with one billion edges.
The number of iterations required in the MapReduce algorithms to find
the SSSP to all vertices for the 3 increasing R-MAT sizes was X, Y, Z
respectively.

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_sssp.pdf}
\caption{Execution times for SSSP using MR-MPI with R-MAT matrices.  
Both the original algorithm (Figure~\ref{fig:sssp}) and the enhanced 
algorithm (Figure~\ref{fig:sssp2}) are included.}
\label{f:sssp}
\end{figure}

To compare our MR-MPI implementation with a wider set of algorithms,
we performed experiments comparing MR-MPI, Hadoop, PBGL (Parallel
Boost Graph Library)~\cite{PBGL} and a multi-threaded implementation
on the Cray XMT using two web graphs: {WebGraphA} with 13.2M vertices
and 31.9M edges, and {WebGraphB} with 187.6M vertices and 531.9M
edges.  In Figure~\ref{f:ssspA}, we show execution times for the SSSP
algorithm using {WebGraphA}.  Like our MR-MPI implementation, the
Hadoop implementation is a Bellman-Ford-style~\cite{Bellman58,Ford62}
algorithm.  The XMT and PBGL implementations are based on
delta-stepping~\cite{MeyerSanders98}, and do not require full
iterations over the entire edge list to advance a breadth-first
search.  We observe that the MR-MPI implementation runs in less time
than the Hadoop implementation, but requires significantly more time
than the XMT and PBGL implementations.  In experiments with
{WebGraphB}, the benefit of the enhanced algorithm
(Figure~\ref{fig:sssp2}) is clearly shown, with a 40\% reduction in
execution time compared to the original SSSP algorithm
(Figure~\ref{fig:sssp}).  But the Bellman-Ford-style iterations are
especially costly in the MR-MPI and Hadoop implementations for
{WebGraphB}, which required 110 iterations to complete; execution
times for this data set are shown in Table~\ref{t:ssspB}.

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_ssspA.pdf}
\caption{SSSP using MR-MPI for WebGraphA with
13.2M vertices and 31.9M edges.  Runtimes using Hadoop and PBGL
are also shown.}
\label{f:ssspA}
\end{figure}

\begin{table}
\begin{center}
\begin{tabular}{|l|c|r|}
\hline
Implementation & Number of Processes & SSSP Execution Time \\
\hline
Hadoop & 48  & 38,925 secs.\\
MR-MPI original (Figure~\ref{fig:sssp}) & 48 &  13,505 secs.\\
MR-MPI enhanced (Figure~\ref{fig:sssp2}) & 48 &  8,031 secs.\\
XMT/C  & 32 &  37 secs.\\
%Out-of-core MR-MPI & 96 &   8,358 secs.\\
%Out-of-core MR-MPI & 64 &  12,882 secs.\\
%Out-of-core MR-MPI & 100 &  6,280 secs.\\
\hline
\end{tabular}
\caption{Execution times for SSSP with {WebGraphB}.}
\label{t:ssspB}
\end{center}
\end{table}

\subsection{Scalability to large numbers of processors}

Finally, we demonstrate the scalability of our MR-MPI library to large
numbers of processors.  The library was used on Sandia's Redstorm and
Thunderbird parallel computers.  Redstorm is a large Cray XT3 with 2+
{GHz} dual/quad-core AMD Opteron processors and a custom interconnect
providing 9.6 GB/s of interprocessor bandwidth.  Thunderbird is a
large Linux cluster with 3.6 {GHz} dual-core Intel EM64T processors
connected by an Infiniband network.  Because these systems do not have
local disks for each processor, we selected a data set and page sizes
that fit in memory, so out-of-core operations were not needed.  For
these experiments, we used two R-MAT data sets with with $2^{25}$
vertices and $2^{28}$ edges, with parameters given in
Table~\ref{t:rmat}.  The ``nice'' parameters produce a less-skewed
sparsity pattern in the matrix than the ``nasty'' parameters, as
indicated by the maximum degree metric.  We ran both the PageRank and
Connected Components algorithms on these matrices.

\begin{table}
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|c|c|}
\hline
Data & R-MAT  & R-MAT  & R-MAT  & R-MAT  & \# of    & \# of & Maximum \\
Set  & a      & b      & c      & d      & vertices & edges & vertex degree\\
\hline
nice  & 0.45 & 0.15 & 0.15 & 0.25 & $2^{25}$ & $2^{28}$ & 1108 \\
nasty & 0.57 & 0.19 & 0.19 & 0.05 & $2^{25}$ & $2^{28}$ & 230,207\\
\hline
\end{tabular}
\caption{Characteristics of R-MAT input data for PageRank and Connected
Components scalability experiments.}
\label{t:rmat}
\end{center}
\end{table}

Figure~\ref{f:prbig}, shows the performance of the various PageRank
implementations on distributed memory and multi-threaded
architectures.  The MR-MPI and Trilinos implementations are described
in Sections~\ref{subsec:graph_pagerank}
and~\ref{subsec:results_pagerank}, respectively.  In the
multi-threaded MTGL implementation, rank propagates via adjacency list
traversal in a compressed sparse-row data structure.  To maintain its
scalability, code must be written so that a single thread spawns the
loop that processes all in-neighbors of a given vertex; this detail
enables the compiler to generate hotspot-free code.

The MR-MPI implementation demonstrated good scalability up to 1024
processors; however, as before, it required an order-of-magnitude more
execution time than the matrix-based implementations on Redstorm.  The
Trilinos implementation is competitive with the multi-threaded
implementation in MTGL on the Cray XMT, when both are run on the same
number of processors.

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_pagerank_big.pdf}
\caption{Scalability comparison of PageRank using MapReduce (MR-MPI),
matrix-based (Trilinos), and multi-threaded (MTGL) implementation on
the R-MAT data sets in Table~\ref{t:rmat}.}
\label{f:prbig}
\end{figure}

Similar results were obtained for the Connected Components algorithm,
as shown in Figure~\ref{f:ccbig}.  As with PageRank, the MR-MPI
implementation showed good scalability up to 1024 processors, but
required significantly more time than the hybrid algorithm using
Trilinos and MR-MPI or the MTGL algorithm.

\begin{figure}[htb]
\includegraphics[width=\textwidth]{fig_cc_big.pdf}
\caption{Scalability comparison of Connected Components algorithms using 
MapReduce (MR-MPI),
matrix-based/MapReduce hybrid (Trilinos/MR-MPI), and MTGL implementations
on the R-MAT data sets in Table~\ref{t:rmat}.}
\label{f:ccbig}
\end{figure}